---
title: "Il1r2 expression and Tregs including prolif Tregs "
output: html_notebook
---

## 1. Header and Environment Setup
```{r}
# --- STEP 1: Load Environment and Global Variables ---
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(Seurat))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(clustree))
 suppressPackageStartupMessages(library(ggalluvial))
```
```{r}
# Define output paths and specific naming conventions for the analysis run
path.output='/Users/iha-eun/Library/CloudStorage/OneDrive-VrijeUniversiteitBrussel/Useful/Own analysis/results'
dataset.name="DL37-38"
subset_name="Tregs_incl.prolif"

# Import custom helper functions for QC and Normalization
source("~/Documents/scripts/preprosessing_May.2020.R")

# Define a consistent color palette for cell clusters in plots
cols.use=c("brown1","goldenrod1","darkorchid","darkolivegreen1","magenta","goldenrod3","dodgerblue","mediumpurple","turquoise3","lightblue","seagreen","coral","violetred","grey","green3","goldenrod4","palevioletred","yellow", "brown3", "pink", "bisque", "yellowgreen", "purple","lightgreen","royalblue","pink3","lightgreen","violet","coral","violetred","grey","green4","goldenrod4")
```

## 2. Data Integration and External Reference
Use these when loading your Seurat object and external Differential Expression (DE) results.
```{r,warning=FALSE, echo=FALSE}
# --- STEP 2: Loading Comparative Reference Data ---
# Import Differential Expression (DE) results from the Zhang 2020 study (MC38 tumor model)
# We use this as a 'ground truth' to see if our Tregs match known tumor-infiltrating signatures.
DEgenes_MC38_Tregs <- readRDS("/Users/iha-eun/Library/CloudStorage/OneDrive-VrijeUniversiteitBrussel/Useful/Own analysis/DEgenes_control.only_Tregs_ZhangL2020_mouse_MC38_tumor_Tcells_annot.rds")

# Load the core Seurat object containing the Treg population
MC38_Tregs<-readRDS(paste0("/Users/iha-eun/Library/CloudStorage/OneDrive-VrijeUniversiteitBrussel/Useful/Own analysis/control.only_Tregs_ZhangL2020_mouse_MC38_tumor_Tcells.seurat.rds"))
```


```{r}
seur_full=readRDS( paste0(path.output,dataset.name,".seurat.rds"))
subset_old=readRDS( paste0(path.output,"Tregs_",dataset.name,".seurat.rds"))
```

```{r}
# --- Global Population Mapping ---
# Define a comprehensive color palette for major immune lineages (T cells, NK, B, Myeloid).
cols.use2=c("Treg"="brown1",  "Treg prolif"="brown3","CD4T"="orange", "CD8T"="yellow2" ,"T prolif"="darkolivegreen1" ,  "NK"="mediumpurple" ,"B"= "magenta"   , "Neutrophil"="cyan"   , "Mono"="lightblue" ,      "Mac"="dodgerblue" ,"cDC2"="green3" ,"pDC"="lightgreen","cDC1"="green4" ,  "migDC"="pink"     )
```


```{r , fig.height = 6, fig.width =7,warning=FALSE}
# Visualize the full dataset to ensure all major immune cell types are present.
DimPlot(seur_full, group.by =  "annot",label=T,repel=T, cols=cols.use2) 
```

```{r , fig.height = 4, fig.width =13}
# --- Multi-Lineage Validation (Marker DotPlot) ---
# Verify cluster identities using canonical markers:
# T-cells (Cd3e), Tregs (Foxp3/Il2ra), Proliferation (Mki67), 
# B-cells (Cd19), and various Myeloid subsets (Lyz2, H2-Aa).
DotPlot(seur_full, group.by = "annot",features=c("Cd3e","Cd3g","Foxp3","Il2ra","Tnfrsf9","Ikzf2","Izumo1r","Mki67","Stmn1","Cd4","Cd40lg","S1pr1","Cd8a","Cd8b1","Xcl1","Klra7","Klrb1c","Ncr1","Gzma","Cd79a","Cd79b","Cd19","Ms4a1","Csf3r","Retnlg","S100a8","Lyz2","Fn1","Chil3","F13a1","Vcan","Tgfbi","Adgre1","C1qa","Ms4a7","Lgmn","H2-Aa","Cd74","H2-DMb1","Cd209a","Napsa","Kmo","Ccr9","Siglech","Pacsin1","Xcr1","Rab7b","Clec9a","Flt3","Cacnb3","Ccl22","Nudt17"  ))+RotatedAxis()+theme(axis.title=element_blank())
```


```{r , fig.height = 5, fig.width = 6}
# --- Target Gene Inspection: Il1r2 ---
# Check the distribution of Il1r2 across the whole dataset using Feature and Violin plots.
# This confirms whether Il1r2 is specifically enriched in Tregs compared to other immune cells.
FeaturePlot(seur_full,c("Il1r2"))
```

```{r , fig.height = 4, fig.width = 6}
VlnPlot(seur_full,group.by = "annot",c("Il1r2"), cols=cols.use, pt.size = 0.2)+NoLegend()+theme(axis.title.x = element_blank())
```


```{r , fig.height =4, fig.width =4}
DotPlot(seur_full,group.by = "annot",features=c("Il1r2"))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r , fig.height = 4, fig.width =6,warning=FALSE}
DimPlot(subset_old, pt.size = 1,label=T,repel=T) 
```


```{r}
# --- Population Refinement and Merging ---
# Isolate "Treg prolif" cells from the full dataset and merge them with the existing Treg subset.
# This ensures the subsequent analysis includes both resting and dividing Treg populations.
Idents(seur_full)=seur_full$annot
subset=merge(subset_old, seur_full[,WhichCells(seur_full, idents="Treg prolif")])

# Preserve the original labels as 'annot.old' before starting new sub-clustering.
subset$annot.old=Idents(subset)
```


```{r,warning=FALSE}
# Check the dimensions of the new subset to confirm successful merging.
dim(subset)
```



## 3. Feature Selection & Dimensionality Reduction

### Preprocessing with removing mitochondrial and roibosomal genes from the HVG list


```{r}
# --- STEP 3: Filtering Variable Features & PCA ---
# Identify mitochondrial (mt-) and ribosomal (Rps/Rpl) genes. 
# These are often excluded from downstream analysis because their high variance 
# can represent cellular stress or technical noise rather than cell-type biology.
mito.genes <- grep(pattern = "^mt-", x = rownames(subset), value = TRUE)
print(mito.genes)
```
```{r}
ribo.genes <- grep(pattern = "^Rps", x = rownames(subset), value = TRUE)
ribo.genes <- c(ribo.genes,grep(pattern = "^Rpl", x = rownames(subset), value = TRUE))
print(ribo.genes)
```

```{r}
# Calculate the percentage of ribosomal reads per cell to monitor protein synthesis levels.
subset$perc.ribo <- Matrix::colSums(GetAssayData(subset, slot = "counts")[ribo.genes, ])/Matrix::colSums(GetAssayData(subset, slot = "counts"))
```

```{r}
mito.genes <- grep(pattern = "^mt-", x = rownames(subset), value = TRUE)
print(mito.genes)
```


```{r , fig.height =4, fig.width = 12,warning=FALSE}
# Visualize 'contamination' or metabolic activity on the UMAP to ensure clusters 
# aren't forming solely based on high mitochondrial/ribosomal content.
FeaturePlot(object = subset, features = c("perc.mito","perc.ribo"))
```
```{r}
# --- Refining Highly Variable Genes (HVGs) ---
# Extract the standard HVG list and manually remove the mito/ribo genes identified above.
# This forces the Principal Component Analysis (PCA) to focus on functional biological markers.
HVG =VariableFeatures(subset)
length(HVG)
```
```{r}
HVG =HVG[!HVG %in% c(ribo.genes,mito.genes)]
length(HVG)
```


```{r,warning=FALSE}
# Run PCA using the cleaned HVG list to reduce the complexity of the dataset.
subset <- RunPCA(subset, features = HVG,  reduction.name = "PcaHVG", reduction.key ="PcaHVG_", verbose=F)
```
```{r , fig.height = 4, fig.width = 6,warning=FALSE}
# --- Determining Dimensionality ---
# Use an Elbow Plot to identify the 'elbow' point where adding more PCs 
# provides diminishing returns in explaining the variance of the data.
ElbowPlot(object = subset,ndims =50,reduction = "PcaHVG")
```

```{r , fig.height = 12, fig.width = 12}
# Use Heatmaps to visualize the primary genes driving each PC. 
# This helps confirm that the selected PCs (1:18) represent real biological signatures.
DimHeatmap(subset, dims = 1:21, balanced = TRUE,reduction = "PcaHVG")
```



```{r,warning=FALSE}
# --- Final Manifold Embedding (UMAP) ---
# Set the number of PCs (18) based on the Elbow Plot and Heatmaps.
dims.use.subset<-18
```



```{r}
# Generate the UMAP coordinates using the PcaHVG reduction for final visualization.
subset <- RunUMAP(subset, dims = 1:dims.use.subset, reduction = "PcaHVG" ,verbose =F, n.neighbors = 20, reduction.name = "umapHVG", reduction.key ="umapHVG_") 
```

```{r , fig.height =3, fig.width = 4,warning=FALSE, echo=FALSE}
# --- Biological Validation of the Embedding ---
# Check how cells group by experimental condition and original annotation.
DimPlot(object = subset,group.by="condition",reduction = "umapHVG", label=T, repel=T)
```

```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, group.by = "annot.old",reduction = "umapHVG",  label=T, repel=T, cols=cols.use)
```
```{r , fig.height = 4, fig.width = 8,warning=FALSE}
# Overlay target gene expression (Il1r2, Bmyc, and Interferon-Stimulated Genes like Isg15) 
# to see where specifically 'fragile' or activated Tregs reside on the map.
FeaturePlot(object = subset,  features = c("Il1r2","Bmyc" ), ncol=2,reduction = "umapHVG")
```
```{r , fig.height = 4, fig.width = 12,warning=FALSE}
FeaturePlot(object = subset,  features = c("Ifit3","Ifit3b", "Isg15" ), ncol=3,reduction = "umapHVG")
```

## 4. Unsupervised Clustering & Marker Discovery

#### Clustering
```{r ,warning=FALSE}
# --- STEP 4: Clustering & Resolution Optimization ---
# Build a Shared Nearest Neighbor (SNN) graph based on the 18 PCs selected earlier.
# The 'k.param=10' determines how many 'neighbors' each cell looks at to find its group.
subset <- FindNeighbors(subset, dims = 1:dims.use.subset, reduction = "PcaHVG", verbose=F, k.param=10, graph.name = "RNA_snn_HVG")

# Run a loop to test clustering at different 'resolutions' (from 0 to 2).
# Algorithm 4 (Leiden) is used here as it is mathematically more robust than the standard Louvain.
for ( i in seq(0,2, 0.25))
subset <- FindClusters(subset, graph.name = "RNA_snn_HVG", resolution = i, verbose=F, algorithm = 4)
# algorithm= 4 is Leiden algorithm - often performs better
```



#### Plot of a clustering tree showing the relationship between clusterings at different resolutions. (using the clustree package)
```{r , fig.height = 8, fig.width = 8,warning=FALSE}
# --- Visualizing Cluster Stability ---
# Use 'clustree' to see how clusters split and merge as resolution increases.
# This helps identify 'stable' clusters versus those that are being forced apart artificially.
clustree(subset, prefix = "RNA_snn_HVG_res.")+
   ggtitle("Clustering tree")
```


```{r , fig.height = 6, fig.width =8,warning=FALSE}
plot<-list()
for ( res in c( 0.75,1,1.25,1.5))
plot[[as.character(res)]]<-DimPlot(subset, pt.size = 1,label=T,repel=T, group.by = paste0("RNA_snn_HVG_res.",res), cols=cols.use,reduction = "umapHVG") +
   ggtitle(paste("res=",res))
# Compare UMAPs of the most likely resolutions (0.75 to 1.5).
# This visual check ensures the chosen resolution captures the biological diversity 
# (e.g., separating Proliferating Tregs from Interferon-high Tregs).
plot_grid(plotlist=plot)
```


```{r ,warning=FALSE}
# --- Final Cluster Selection ---
# Set the final resolution to 1.0 and assign these labels as the active identities (Idents).
res=1
Idents(subset)=paste0("RNA_snn_HVG_res.",res)
```
```{r , fig.height = 3, fig.width =4,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use,reduction = "umapHVG") 
```

```{r , fig.height =3, fig.width =5}
# Final check: Do these clusters express our target gene?
# Plotting Il1r2 expression across the newly defined clusters.
DotPlot(subset, features = c( "Il1r2"), group.by = paste0("RNA_snn_HVG_res.",1))+RotatedAxis()
```

```{r}
# --- Automated Marker Identification (All Clusters) ---
# Loop through every cluster to find 'Marker Genes' (genes significantly higher in one cluster 
# compared to all others combined).
DEgenes=list()
for (i in levels(Idents(subset))){
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = i,min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
}

# Save the list of marker genes. These will be used to 'name' the clusters 
# (e.g., if a cluster has Mki67 as a top marker, we name it 'Treg prolif').
saveRDS(DEgenes, paste0( path.output,"DEgenes_HVGrm_PC",dims.use.subset,".res",res,"_", subset_name,"_",dataset.name,".rds"))
```


```{r}
res=1
dims.use.subset=18
DEgenes <-readRDS( paste0( path.output,"DEgenes_HVGrm_PC",dims.use.subset,".res",res,"_", subset_name,"_",dataset.name,".rds"))
```

## 5. Biological Validation & Sub-Population Profiling

#### Check the expression of the top DE genes using a DotPlot
```{r}
# --- STEP 5A: Top Marker Identification ---
# Extract the top 8 differentially expressed (DE) genes for each cluster to create a 
# 'genetic fingerprint' for every sub-population.
features.use=unlist(lapply(DEgenes, function(x) { head(x[x$avg_logFC>0,]$gene,8)}))
length(features.use)
names(features.use)=NULL
features.use=features.use[!duplicated(features.use)]
```

```{r , fig.height =4, fig.width =14}
# Broad DotPlot: Visualize these top markers to define cluster identities 
# (e.g., finding the Mki67+ proliferating cluster or the Isg15+ interferon cluster).
DotPlot(subset, features = features.use)+RotatedAxis()
```
```{r , fig.height = 3, fig.width =4,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use,reduction = "umapHVG") 
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c("Cd4","Cd40lg","Foxp3","Icos","Ccr2","Ly6a","Maf" ,"Il2ra","Ctla4" ,"Bcl2a1b","Ccl4","Nr4a1","Nr4a3" ,"Csf1" ,"Il10","Ccr8","Il1r2","Cst7" ,"Chchd10", "Tnfrsf18","S100a4" ,"Tff1","Bbx","Ccdc184" , "Eno1" ,"Bnip3",  "Hilpda" , "Ddit3"   ,"Bmyc","Npm1","Serbp1", "Ifngr1" ,"Il7r",  "Gimap1" ,"Ctss","Lef1" , "Ifi203","Ifit3","Isg15" ,"Ran","Mki67","Stmn1","doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```



#### Standard preprocessing and clustering of the subsetted dataset
```{r,warning=FALSE}
# --- STEP 5B: Standardized Re-processing ---
# To ensure the highest resolution for this specific Treg subset, 
# re-normalize and re-scale the data, then recalculate Principal Components.
subset <-  NormalizeData(subset,verbose = F) %>% FindVariableFeatures(verbose = F) %>% ScaleData(verbose = F) 
subset=RunPCA(subset, features = VariableFeatures(subset) ,verbose = F)
```
```{r , fig.height = 4, fig.width = 6,warning=FALSE}
# Elbow Plot: Determining the optimal dimensions (setting dims to 12 based on the drop-off).
ElbowPlot(object = subset,ndims =50)
```
```{r , fig.height = 12, fig.width = 12}
# Heatmaps: Identifying which biological programs (e.g., cell cycle vs signaling) 
# drive each Principal Component (PC).
DimHeatmap(subset, dims = 1:20, balanced = TRUE)
```

```{r , fig.height = 12, fig.width = 12}
DimHeatmap(subset, dims = 21:40, balanced = TRUE)
```

```{r,warning=FALSE}
dims.use.subset<-12
```

```{r,warning=FALSE}
table(Idents(subset))
```

```{r}
subset <- RunUMAP(subset, dims = 1:dims.use.subset,verbose =F, n.neighbors = 20) 
```

```{r , fig.height =3, fig.width = 4,warning=FALSE, echo=FALSE}
DimPlot(object = subset,group.by="condition", label=T, repel=T)
```

```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, label=T, repel=T, cols=cols.use)
```

```{r , fig.height =3, fig.width =12}
# --- STEP 5C: Master Annotation DotPlot (The Biological 'ID Card') ---
# This massive plot validates every cluster using four major gene 'modules':
# - Lineage: Cd4, Foxp3, Il2ra (Ensures they are all Tregs)
# - Activation/Suppression: Ctla4, Il10, Il1r2, Ccr8
# - Fragility/Interferon: Ifngr1, Ifit3, Isg15, Ifi203
# - Proliferation: Mki67, Stmn1, Mcm3
# - Quality Control: doublet.score, perc.mito (Ensures no bad cells are driving a cluster)
DotPlot(subset, features = c("Cd3e","Cd3g","Cd4","adt_CD4","Foxp3","Il2ra","Ctla4","Lgals3bp" , "Arl5a" ,"Ifi203" ,"Cst7" ,"Il10","Med20","Tmem126b","Il1r2","Ccr8","Rpl28",     "Rpl22l1" ,"Ran","Tff1","Dnlz",  "Ifngr1","Neurl3", "Shisa5" ,  "Gimap1" ,"Ctss","Ifit3","Ifit3b", "Hilpda" , "Ddit3"   , "Eno1" ,"Bnip3" ,"Serpina3f","Sell","adt_CD62L","Mki67","Stmn1","Mcm3","doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```



 
```{r}
# --- STEP 5D: External Validation ---
# Using the Zhang 2020 (MC38) reference list, extract genes associated with 'Ctla4' 
# and project them onto this dataset to see if the effector signatures match.
i="Treg.Ctla4"  
selected.genes= DEgenes_MC38_Tregs[[i]][ DEgenes_MC38_Tregs[[i]]$p_val_adj<1e-3 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.5,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c(selected.genes))+RotatedAxis()+theme(axis.title = element_blank())
```

## 6. Mapping Treg Functional "Flavors" via Spatial Expression

Il1r2-lo Ctla4-hi
```{r , fig.height = 6, fig.width = 12,warning=FALSE}
# --- STEP 6: Spatial Validation of Treg Subpopulations ---
# We use FeaturePlots to project gene expression onto the UMAP coordinates. 
# This confirms which mathematical clusters correspond to specific biological states.

# --- 6A: Identifying 'Classical' Effector Tregs (Il1r2-lo / Ctla4-hi) ---
# Characterized by high Ctla4 and Il10 (suppression) but lower Il1r2. 
# These represent the 'stable' activated Treg pool.
FeaturePlot(object = subset,  features = c("Ctla4","Lgals3bp" , "Arl5a"  ,"Cst7" ,"Il10","Med20" ), ncol=3)
```

```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, label=T, repel=T, cols=cols.use)
```

Il1r2-hi
```{r , fig.height = 8, fig.width = 12,warning=FALSE}
# --- 6B: Identifying the Primary Target (Il1r2-hi / Activated) ---
# Mapping our main population of interest. Note the overlap with 
# proliferation markers (Mki67, Stmn1) and the terminal activation marker Ccr8.
FeaturePlot(object = subset,  features = c("Il1r2","Ccr8","Rpl28",     "Rpl22l1" ,"Ran","Tff1","Dnlz" , "Mki67","Stmn1"), ncol=3)
```

```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, label=T, repel=T, cols=cols.use)
```


Ifngr1-hi
```{r , fig.height = 8, fig.width = 12,warning=FALSE}
# --- 6C: Mapping the Interferon-Response / 'Fragile' State (Ifngr1-hi) ---
# High expression of Ifit3 and Ifit3b indicates an interferon-stimulated state, 
# which is a hallmark of Treg 'fragility' and loss of suppressive stability.
FeaturePlot(object = subset,  features = c("Ifngr1","Neurl3", "Shisa5" ,  "Gimap1" ,"Ctss","Ifit3","Ifit3b" ), ncol=3)
```

```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, label=T, repel=T, cols=cols.use)
```

Hypoxic
```{r , fig.height = 8, fig.width = 12,warning=FALSE}
# --- 6D: Mapping the Hypoxic / Metabolic Stress State ---
# Cells expressing Hilpda, Eno1, and Bnip3 are likely located in the necrotic 
# or low-oxygen regions of the tumor, shifting toward glycolysis.
FeaturePlot(object = subset,  features = c("Hilpda" , "Ddit3"   , "Eno1" ,"Bnip3" ,"Serpina3f" ), ncol=3)
```

```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, label=T, repel=T, cols=cols.use)
```
Il1r2-hi Ccr8-hi
```{r , fig.height = 8, fig.width = 12,warning=FALSE}
# --- 6E: Distinguishing Specialized Il1r2-hi Sub-states ---
# Comparing two distinct activated flavors:
# 1. Ccr8-hi/Il10-hi: Highly suppressive, terminally differentiated tumor Tregs.
# 2. Bmyc-hi/Cd83-hi: A specific sub-state associated with activation and CD70 signaling.
FeaturePlot(object = subset,  features = c("Ccr8",  "Ccl4" ,"Il10" , "Bcl2a1d" , "Vsir" ,"Irf8", "Glipr1"), ncol=3)
```

Il1r2-hi Bmyc-hi
```{r , fig.height = 8, fig.width = 12,warning=FALSE}
FeaturePlot(object = subset,  features = c( "Bmyc" ,"Cd83"  ,  "Nrip1" , "Chchd10" , "Gda","Hspa9"  ,  "Dgat2" ,"Cd70", "Gadd45g"   ), ncol=3)
```


## 7. Resolution Optimization and Signature Cross-Validation

#### Clustering
```{r ,warning=FALSE}
# --- STEP 7A: Iterative Clustering & Stability Testing ---
# Build a Shared Nearest Neighbor (SNN) graph. Using 'k.param=15' increases 
# the local connectivity of the cells for more stable cluster boundaries.
subset <- FindNeighbors(subset, dims = 1:dims.use.subset, verbose=F, k.param=15)

# Test a broad range of resolutions (0 to 2) using the Leiden algorithm (Algorithm 4).
# Higher resolutions allow us to catch small but biologically significant sub-populations.
for ( i in seq(0,2, 0.25))
subset <- FindClusters(subset, resolution = i, verbose=F, algorithm = 4)
# algorithm= 4 is Leiden algorithm - often performs better
```



#### Plot of a clustering tree showing the relationship between clusterings at different resolutions. (using the clustree package)
```{r , fig.height = 8, fig.width = 8,warning=FALSE}
# --- STEP 7B: Visualizing Cluster Relationships ---
# Plot a 'Clustree' to determine at which point clusters become unstable or start 
# swapping cells. This justifies selecting resolution 1.5 as the final setting.
clustree(subset, prefix = "RNA_snn_res.")+
   ggtitle("Clustering tree")
```


```{r , fig.height = 6, fig.width =8,warning=FALSE}
plot<-list()
for ( res in c( 0.75,1,1.25,1.5))
plot[[as.character(res)]]<-DimPlot(subset, pt.size = 1,label=T,repel=T, group.by = paste0("RNA_snn_res.",res), cols=cols.use) +
   ggtitle(paste("res=",res))

# Compare the spatial arrangement of clusters across different high resolutions.
plot_grid(plotlist=plot)
```


```{r ,warning=FALSE}
# Final Identity Assignment: Setting the active idents to the high-resolution clusters.
res=1.5
Idents(subset)=paste0("RNA_snn_res.",res)
```
```{r , fig.height = 3, fig.width =4,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use) 
```




```{r}
# --- STEP 7C: Cross-Dataset Validation (MC38 Reference) ---
# We validate our clusters by checking if they express signature genes from 
# external tumor-infiltrating Treg (tiTreg) studies:
# 1. 'Il1r2-hi Bmyc-hi' profile (Discovery Validation)
# 2. 'Il1r2-hi Ccr8-hi' profile (Terminal Activation Validation)
# 3. 'Ctla4' & 'Socs2' profiles (Effector/Resting state check)
i="Treg.Il1r2.hi.Bmyc.hi"
selected.genes= DEgenes_MC38_Tregs [[i]][ DEgenes_MC38_Tregs[[i]]$p_val<0.01 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.35,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```



```{r}
i= "Treg.Il1r2.hi.Ccr8.hi"
selected.genes= DEgenes_MC38_Tregs[[i]][ DEgenes_MC38_Tregs[[i]]$p_val<0.01 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.3,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r}
i="Treg.Ctla4"  
selected.genes= DEgenes_MC38_Tregs[[i]][ DEgenes_MC38_Tregs[[i]]$p_val_adj<1e-3 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.5,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c(selected.genes))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r}
i="Treg.Socs2" 
selected.genes= DEgenes_MC38_Tregs[[i]][ DEgenes_MC38_Tregs[[i]]$p_val<1e-2 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.2,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c(selected.genes))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r}
# --- STEP 7D: Detailed Differential Expression (DE) Testing ---
# Automated marker identification for all high-resolution clusters.
DEgenes=list()
for (i in levels(Idents(subset))){
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = i,min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
}
saveRDS(DEgenes, paste0( path.output,"DEgenes_PC",dims.use.subset,".res",res,"_", subset_name,"_",dataset.name,".rds"))
```

```{r}
# --- STEP 7E: Pairwise Cluster Comparison (7 vs 3) ---
# Specifically investigating the differences between two clusters (e.g., 7 and 3). 
# This is often used to compare a 'fragile' cluster vs. an 'effector' cluster.
i="7.vs.3"
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = 7, ident.2 = 3,min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
saveRDS(DEgenes, paste0( path.output,"DEgenes_PC",dims.use.subset,".res",res,"_", subset_name,"_",dataset.name,".rds"))
```


```{r , fig.height =4, fig.width =12}
# Volcano Plot: Visualizing the magnitude and significance of the 7 vs 3 comparison.
VolcanoPlot( DEgenes[[i]], logFCmin_neg_col = -0.5)+ggtitle(i)
```
```{r}
# Save the finalized DE results for the chosen resolution and PC set.
res=1.5
dims.use.subset=12
DEgenes <-readRDS( paste0( path.output,"DEgenes_PC",dims.use.subset,".res",res,"_", subset_name,"_",dataset.name,".rds"))
```


#### Check the expression of the top DE genes using a DotPlot
```{r}
features.use=unlist(lapply(DEgenes, function(x) { head(x[x$avg_logFC>0,]$gene,8)}))
length(features.use)
names(features.use)=NULL
features.use=features.use[!duplicated(features.use)]
```

```{r , fig.height =4, fig.width =14}
DotPlot(subset, features = features.use)+RotatedAxis()
```
```{r , fig.height = 3, fig.width =4,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use) 
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c( "Ifngr1", "Shisa5" ,  "Gimap1" ,"Ctss","Maf" ,"Ctla4","Ncor1","Btg1" ,"Crlf2","perc.ribo", "Cd4","Cst7" ,"Bcl2a1b" ,"Csf1","Rel" , "Il2ra","Foxp3","Icos","Ccr2","Ly6a","Il7r","Ccr8","Il1r2","Bmyc" ,"Mif","Slc39a7","Dnajb14","Zfp593","Nhp2","Actr5","Tff1","Myo18a","Scnm1", "Chchd10","Cd83" ,  "Nrip1" ,"Icam1",   "Hilpda" , "Ddit3"   , "Eno1" ,"Bnip3" , "Ifi203","Ifit3","Isg15" ,"Ran","Mki67","Stmn1","doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```

## 8. Biological Annotation and Multi-Level Population Mapping

```{r ,warning=FALSE}
# --- STEP 8A: Defining Detailed Cluster Identities ---
# We map the mathematical cluster numbers to biological names based on their 
# expression of 'marker genes' (e.g., Ifngr1, Ctla4, Il1r2, Mki67).
# Note the identification of specialized subsets like 'Treg IFN' and 'Treg Bmyc-hi'.
Idents(subset)=plyr::mapvalues(Idents(subset), from=1:8, 
    to=c("Treg Ifngr1-hi ribo-hi", "Treg Ifngr1-hi Csf1","Treg Il1r2-hi","Treg Ifngr1-hi Ctla4-hi","Treg IFN",
        "Treg Il1r2-hi prolif","Treg Il1r2-hi Ccr2-hi","Treg  Bmyc-hi"))
```


```{r}
new.order=c ( "Treg Ifngr1-hi ribo-hi","Treg IFN", "Treg Ifngr1-hi Csf1","Treg Ifngr1-hi Ctla4-hi","Treg  Bmyc-hi" ,"Treg Il1r2-hi Ccr2-hi" ,"Treg Il1r2-hi",
        "Treg Il1r2-hi prolif"    )
new.order[!new.order %in% levels( Idents(subset))]
levels( Idents(subset))[!levels( Idents(subset)) %in%new.order]
```



```{r}
Idents(subset)=factor( Idents(subset), levels=new.order)
subset$annot.detailed=Idents(subset)
```




```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use) 
```

```{r , fig.height =4, fig.width =6,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use, group.by = "annot.old") 
```


```{r , fig.height =8, fig.width =10,warning=FALSE}
# --- STEP 8B: Cell Cycle and Proliferation Validation ---
# Use FeaturePlots of Birc5 and Stmn1 to double-check the proliferating cluster.
# CellSelector allows for manual refinement of the 'proliferating' identity.
FeaturePlot(subset,c("Mki67","Birc5","Stmn1","Mcm3")) 
```

```{r , fig.height =4, fig.width =6,warning=FALSE}
DimPlot(subset,label=T,repel=T, group.by = "Phase") 
```

```{r , fig.height =4, fig.width =14}
# --- STEP 8C: Broad Marker Validation (The Final DotPlot) ---
# Validating the new groupings against the full suite of activation (Ctla4, Il1r2), 
# metabolism (Hilpda, Eno1), and interferon (Ifit3, Isg15) genes.
DotPlot(subset, group.by = "annot.detailed", features = c( "Ifngr1", "Crlf2","perc.ribo", "Shisa5"  ,"Ctss","Ifi203","Ifit3","Isg15" ,"Ifng","Cd4","Cst7" ,"Bcl2a1b" ,"Csf1","Rel" ,"Maf" ,"Ctla4","Ncor1","Btg1" , "Bmyc" ,"Mif","Chchd10", "Nrip1" ,"Icam1",   "Hilpda" ,"Cd83" ,  "Ddit3"   , "Bnip3" ,"Eno1" ,"Il2ra","Foxp3","Icos","Ccr2","Ly6a","Il7r","Ccr8","Il1r2","Slc39a7","Dnajb14","Zfp593","Nhp2","Actr5","Tff1","Myo18a","Scnm1",  "Ran","Mki67","Stmn1","doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```
```{r , fig.height =3, fig.width =5}
DotPlot(subset, group.by = "annot.detailed", features = c( "Il1r2"))+RotatedAxis()
```


```{r , fig.height =3, fig.width =5}
DotPlot(subset, group.by = "annot.old", features = c( "Il1r2"))+RotatedAxis()
```

```{r , fig.height =3, fig.width =5}
DotPlot(subset, group.by = "annot", features = c( "Il1r2"))+RotatedAxis()
```



```{r}
# --- STEP 8D: Creating Hierarchical Metadata ---
# We create different 'levels' of granularity for the analysis:
# 1. annot.detailed: Every sub-cluster kept separate.
# 2. annot.general: Grouping all Ifngr1-hi/Ctla4-hi cells into a broad 'Treg' category.
# 3. annot: A middle-ground grouping for balanced statistical comparisons.
subset$annot=plyr::mapvalues(Idents(subset), from=c("Treg Ifngr1-hi ribo-hi", "Treg Ifngr1-hi Csf1","Treg Ifngr1-hi Ctla4-hi","Treg  Bmyc-hi" ), to=c("Treg", "Treg","Treg","Treg" ))
Idents(subset)="annot"
Idents(subset, cells=WhichCells(subset, expression= annot.old=="Treg prolif"))="Treg Il1r2-hi prolif"
subset$annot=Idents(subset)
subset$annot=factor( subset$annot, levels=c("Treg", "Treg IFN"   ,  "Treg Il1r2-hi Ccr2-hi", "Treg Il1r2-hi" ,"Treg Il1r2-hi prolif"))
```





```{r , fig.height =8, fig.width =8,warning=FALSE}
FeaturePlot(subset,c("Mki67","Birc5","Stmn1","Mcm3")) 
```


select.cells=CellSelector(FeaturePlot(subset,c("Birc5")))
Idents(subset, cells=select.cells)="Treg Il1r2-hi prolif"
select.cells=CellSelector(FeaturePlot(subset,c("Stmn1")))
Idents(subset, cells=select.cells)="Treg Il1r2-hi prolif"

subset$annot=Idents(subset)
subset$annot=factor( subset$annot, levels=c("Treg", "Treg IFN"   ,  "Treg Il1r2-hi Ccr2-hi", "Treg Il1r2-hi" ,"Treg Il1r2-hi prolif"))

```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot", features = c("Ctla4","Btg1" , "Ifngr1", "Ifi203","Ifit3","Isg15","Il1r2", "Il2ra","Foxp3","Icos","Ccr2","Ly6a","Il7r","Il2rg","Fam120a","Rab27a","Furin" ,"Slfn9","perc.ribo", "Bmyc","Slc39a7","Zfp593","C1qbp","Nhp2","Dnajb14","Tff1","Il17ra","Scnm1","Bnip3" ,"Eno1", "Mif", "Ran","Mki67","Stmn1","Ccr8"))+RotatedAxis()
```
```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use, group.by = "annot") 
```



```{r}
subset$annot.general=plyr::mapvalues(Idents(subset), from=c("Treg Ifngr1-hi ribo-hi", "Treg Ifngr1-hi Csf1","Treg Ifngr1-hi Ctla4-hi","Treg  Bmyc-hi" ,"Treg Il1r2-hi Ccr2-hi"  ), to=c("Treg", "Treg","Treg","Treg" ,"Treg Il1r2-hi"))
Idents(subset)="annot.general"
Idents(subset, cells=WhichCells(subset, expression= annot=="Treg Il1r2-hi prolif"))="Treg Il1r2-hi prolif"
subset$annot.general=Idents(subset)
subset$annot.general=factor( subset$annot.general, levels=c("Treg", "Treg IFN"   ,  "Treg Il1r2-hi" ,"Treg Il1r2-hi prolif"))
```
```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use, group.by = "annot.general") 
```
```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot.general", features = c( "Il2ra","Foxp3","Icos","Ctla4","Btg1" , "Ccr2","Crlf2","Il7r","Ifngr1", "Ifi203","Ifit3","Isg15","Il1r2","perc.ribo", "Bmyc","Fam120a","Slc39a7","Rab27a","Furin" ,"Dnajb14","Tff1","Scnm1","Bnip3" ,"Eno1", "Mif", "Ran","Mki67","Stmn1","Ccr8"))+RotatedAxis()
```
```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"), group.by = "annot.general")+RotatedAxis()
```



```{r}
Idents(subset)="annot.detailed"
Idents(subset, cells=WhichCells(subset, expression= annot=="Treg Il1r2-hi prolif"))="Treg Il1r2-hi prolif"
subset$annot.detailed=Idents(subset)
subset$annot.general=factor( subset$annot.general, levels=c ( "Treg Ifngr1-hi ribo-hi","Treg IFN", "Treg Ifngr1-hi Csf1","Treg Ifngr1-hi Ctla4-hi","Treg  Bmyc-hi" ,"Treg Il1r2-hi Ccr2-hi" ,"Treg Il1r2-hi",  "Treg Il1r2-hi prolif"    ))
```
```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use, group.by = "annot.detailed") 
```



```{r,warning=FALSE, echo=FALSE}
# --- STEP 8E: Comprehensive DE Analysis and Data Export ---
# Running FindMarkers loops across all three metadata levels (detailed, general, annot).
# Results are saved as .rds for R and .xlsx for external review.
clusters=levels(subset$annot.detailed)
DEgenes_list<-list()
for ( i in clusters){
DEgenes_list[[i]]<-  FindMarkers(subset, group.by="annot.detailed",ident.1=i,min.cells.group=1, max.cells.per.ident = 1000, pseudocount.use = 0.1)
DEgenes_list[[i]]$gene=rownames(DEgenes_list[[i]])
DEgenes_list[[i]]$cluster.1=i
DEgenes_list[[i]]$pseudocount.use=0.1
DEgenes_list[[i]]$max.cells.per.ident=1000
}
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```




```{r,warning=FALSE, echo=FALSE}
DEgenes_list=readRDS( paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```
```{r,warning=FALSE, echo=FALSE}
names(DEgenes_list)=make.names(names(DEgenes_list))


### Save the differentially expressed genes as an excel table
writexl::write_xlsx(DEgenes_list,paste0(path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.xlsx"))
```


```{r,warning=FALSE, echo=FALSE}
clusters=levels(subset$annot.general)
DEgenes_list<-list()
for ( i in clusters){
DEgenes_list[[i]]<-  FindMarkers(subset, group.by="annot.general",ident.1=i,min.cells.group=1, max.cells.per.ident = 1000, pseudocount.use = 0.1)
DEgenes_list[[i]]$gene=rownames(DEgenes_list[[i]])
DEgenes_list[[i]]$cluster.1=i
DEgenes_list[[i]]$pseudocount.use=0.1
DEgenes_list[[i]]$max.cells.per.ident=1000
}
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.general.rds"))
```




```{r,warning=FALSE, echo=FALSE}
DEgenes_list=readRDS( paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.general.rds"))
```
```{r,warning=FALSE, echo=FALSE}
names(DEgenes_list)=make.names(names(DEgenes_list))
### Save the differentially expressed genes as an excel table
writexl::write_xlsx(DEgenes_list,paste0(path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.general.xlsx"))
```


```{r,warning=FALSE, echo=FALSE}
clusters=levels(subset$annot)
DEgenes_list<-list()
for ( i in clusters){
DEgenes_list[[i]]<-  FindMarkers(subset, group.by="annot",ident.1=i,min.cells.group=1, max.cells.per.ident = 1000, pseudocount.use = 0.1)
DEgenes_list[[i]]$gene=rownames(DEgenes_list[[i]])
DEgenes_list[[i]]$cluster.1=i
DEgenes_list[[i]]$pseudocount.use=0.1
DEgenes_list[[i]]$max.cells.per.ident=1000
}
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```


```{r,warning=FALSE, echo=FALSE}
DEgenes_list=readRDS( paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```



```{r,warning=FALSE, echo=FALSE}
Idents(subset)="annot"
```
```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"), group.by = "annot")+RotatedAxis()
```
```{r}
# --- STEP 8F: Targeted Comparative Analysis ---
# Statistical comparison specifically between the Il1r2-high population 
# and the combined Resting/IFN Treg population.
i="Il1r2-hi.vs.Treg_and_Treg.IFN"
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi", ident.2 = c("Treg","Treg IFN"),min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
  saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```


```{r , fig.height =5, fig.width =12}
# Visualizing the 'Il1r2-hi' signature markers using a Volcano Plot.
i="Il1r2-hi.vs.Treg_and_Treg.IFN"
VolcanoPlot(  DEgenes[[i]], p_valmax_col = 0.5)
```




```{r}
selected.genes=rownames( DEgenes[[i]][ DEgenes[[i]]$p_val<1e-5 &  DEgenes[[i]]$avg_logFC>1,])
length(selected.genes)
```

Export gene names as a csv file for metascape
```{r}
# Exporting the high-confidence gene list for pathway analysis (e.g., Metascape).
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```


```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"/metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```

```{r}
i="Il1r2-hi-Ccr2.vs.Treg_and_Treg.IFN"
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi Ccr2-hi", ident.2 = c("Treg","Treg IFN"),min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
  saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```


```{r , fig.height =5, fig.width =12}
i="Il1r2-hi-Ccr2.vs.Treg_and_Treg.IFN"
VolcanoPlot(  DEgenes[[i]], p_valmax_col = 0.5)
```

```{r}
selected.genes=rownames( DEgenes[[i]][ DEgenes[[i]]$p_val<1e-3 &  DEgenes[[i]]$avg_logFC>0.8,])
length(selected.genes)
```

Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```

## 9. Pathway Integration & Meta-Cluster Differential Expression

```{r}
# --- STEP 9A: Metascape Pathway Integration ---
# Import the enriched Gene Ontology (GO) terms from external Metascape analysis.
# We filter for 'Summary' groups to focus on the most impactful biological pathways.
GO.metascape=readxl::read_xlsx( paste0(path.output,"/metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```




```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
# --- STEP 9B: Pathway-Specific Volcano Plotting ---
# Visualize the DE results while highlighting genes belonging to a specific GO term.
# This proves that the pathway (e.g., 'T cell activation') is statistically significant.
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```



```{r}
# --- STEP 9C: Targeted Comparison (Ccr2-hi Cluster vs. Resting/IFN) ---
# Isolate the specific differences of the 'Ccr2-hi' Treg subset. 
# Ccr2 is often linked to recruitment and specific localization within the tumor.
i="Il1r2-hi_Il1r2-hi_Ccr2.vs.Treg_and_Treg.IFN"
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = c("Treg Il1r2-hi","Treg Il1r2-hi Ccr2-hi"), ident.2 = c("Treg","Treg IFN"),min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
  saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```


```{r , fig.height =5, fig.width =12}
# --- STEP 9D: Meta-Cluster Signature Analysis ---
# Group both 'Il1r2-hi' and 'Il1r2-hi Ccr2-hi' into a single 'Activated Meta-cluster'.
# This reveals the common core signature of highly activated tumor Tregs 
# regardless of their specific sub-state.
i="Il1r2-hi_Il1r2-hi_Ccr2.vs.Treg_and_Treg.IFN"
VolcanoPlot(  DEgenes[[i]], p_valmax_col = 0.5)
```

```{r}
# --- STEP 9E: High-Confidence Gene Selection for Metascape ---
# Filter for genes with high fold change (>1) and extreme significance (p < 1e-4).
# Export these as .csv files to feed back into the pathway discovery pipeline.
selected.genes=rownames( DEgenes[[i]][ DEgenes[[i]]$p_val<1e-4 &  DEgenes[[i]]$avg_logFC>1,])
length(selected.genes)
```

Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```

## 10. Multi-Layered Functional Validation & Specific Cluster Comparisons

```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"/metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```

### Top GO enriched main pathway:
```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
# --- STEP 10A: Biological Pathway Projection ---
# Map the top enriched GO pathways from Metascape back onto the Volcano Plots.
# This visually highlights how specific biological programs (e.g., 'T cell activation') 
# are distributed within your differential expression results.
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```



```{r}
# --- STEP 10B: Intragroup Comparisons (The 'Sub-cluster' Nuance) ---
# We perform 'internal' differential expression to find the unique markers that 
# separate different flavors of activated Tregs:
# 1. Ccr2-hi vs. Standard Il1r2-hi: Identifying markers for chemotaxis/recruitment.
# 2. Proliferating vs. Standard Il1r2-hi: Identifying markers for cell cycle entry.
i="Il1r2-hi-Ccr2.vs.Il1r2-hi"
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi Ccr2-hi", ident.2 = c("Treg Il1r2-hi"),min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
  saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```


```{r , fig.height =5, fig.width =12}
i="Il1r2-hi-Ccr2.vs.Il1r2-hi"
VolcanoPlot(  DEgenes[[i]], p_value ="p_val" )
```


```{r}
i="Il1r2-hi-prolif.vs.Il1r2-hi"
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi prolif", ident.2 = c("Treg Il1r2-hi"),min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
  saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```


```{r , fig.height =5, fig.width =12}
i="Il1r2-hi-prolif.vs.Il1r2-hi"
VolcanoPlot(  DEgenes[[i]], p_value ="p_val" )
```


```{r,warning=FALSE, echo=FALSE}
DEgenes_list=readRDS( paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.rds"))
```
```{r,warning=FALSE, echo=FALSE}
names(DEgenes_list)=make.names(names(DEgenes_list))
### Save the differentially expressed genes as an excel table
writexl::write_xlsx(DEgenes_list,paste0(path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.xlsx"))
```


```{r , fig.height =5, fig.width =7,warning=FALSE}
DimPlot(subset, group.by = "annot.detailed",label=T,repel=T,cols= cols.use) 
```

```{r , fig.height =4, fig.width =14}
# --- STEP 10C: Cross-Referencing with Reference Signatures ---
# Validate the 'Bmyc-hi' cluster by projecting the MC38 Treg signature. 
# This ensures that your detailed annotations (Step 8) remain consistent 
# with established mouse tumor models.
DotPlot(subset, group.by = "annot.detailed", features = c( "Ifngr1", "Crlf2","perc.ribo", "Shisa5"  ,"Ctss","Ifi203","Ifit3","Isg15" ,"Ifng","Cd4","Cst7" ,"Bcl2a1b" ,"Csf1","Rel" ,"Maf" ,"Ctla4","Ncor1","Btg1" , "Bmyc" ,"Mif","Chchd10", "Nrip1" ,"Icam1",   "Hilpda" ,"Cd83" ,  "Ddit3"   , "Bnip3" ,"Eno1" ,"Il2ra","Foxp3","Icos","Ccr2","Ly6a","Il7r","Ccr8","Il1r2","Slc39a7","Dnajb14","Zfp593","Nhp2","Actr5","Tff1","Myo18a","Scnm1",  "Ran","Mki67","Stmn1","doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```


```{r}
# --- STEP 10D: External Reference Validation (MC38 Treg) ---
# Using the Bmyc-hi signature from the MC38 Treg reference study to confirm 
# that our 'annot.detailed' sub-clusters match established biological models.
i="Treg.Il1r2.hi.Bmyc.hi"
selected.genes= DEgenes_MC38_Tregs [[i]][ DEgenes_MC38_Tregs[[i]]$p_val<0.01 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.35,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot.detailed", features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot", features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```




```{r , fig.height =4, fig.width =4.3,warning=FALSE}
FeaturePlot(subset,"Il1r2") 
```

```{r , fig.height =4, fig.width =4,warning=FALSE}

VlnPlot(subset,"Il1r2",cols= cols.use, group.by = "annot.general")+NoLegend() +theme(axis.title.x = element_blank())
```

```{r , fig.height =4, fig.width =6,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use, group.by = "annot.general") 
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot.general", features = c( "Il2ra","Foxp3","Icos","Ctla4","Btg1" , "Ccr2","Crlf2","Il7r","Ifngr1", "Ifi203","Ifit3","Isg15","Il1r2","perc.ribo", "Bmyc","Fam120a","Slc39a7","Rab27a","Furin" ,"Dnajb14","Tff1","Scnm1","Bnip3" ,"Eno1", "Mif", "Ran","Mki67","Stmn1","Ccr8"))+RotatedAxis()
```



```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"), group.by = "annot.general")+RotatedAxis()
```


```{r , fig.height =4, fig.width =4,warning=FALSE}
VlnPlot(subset,"Il1r2",cols= cols.use, group.by = "annot")+NoLegend() +theme(axis.title.x = element_blank())
```

```{r , fig.height =4, fig.width =6,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use, group.by = "annot") 
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot", features = c("Ctla4","Btg1" ,"Ltb", "Ifngr1", "Ifi203","Ifit3","Isg15","Il1r2", "Il2ra","Foxp3","Icos","Ccr2","Ly6a","Il7r","Il2rg","Fam120a","Rab27a","Furin" ,"Slfn9","perc.ribo", "Bmyc","Slc39a7","Zfp593","C1qbp","Nhp2","Dnajb14","Tff1","Il17ra","Scnm1","Bnip3" ,"Eno1", "Mif", "Ran","Mki67","Stmn1","Ccr8"))+RotatedAxis()
```


```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"), group.by = "annot")+RotatedAxis()
```


hypoxia
```{r , fig.height =3, fig.width =13}
# --- STEP 10E: Metabolic & Hypoxia Characterization ---
# Mapping the 'Hypoxia' module (Hilpda, Vegfa, Eno1). This identifies which 
# Tregs are responding to the low-oxygen tumor core environment.
DotPlot(subset, features = c("Hilpda", "Vegfa", "Bnip3", "Hmox1", "Fth1", "Aldoa", "Mif", "Ftl1", "Arg1", "Ldha", "Adam8", "Eno1", "Srgn", "Ndrg1", "Mt1", "Fam162a", "Slc7a2", "Slc2a1", "Mgarp", "Lgals3","Spp1" ))+RotatedAxis()+theme(axis.title = element_blank())
```

Treg activation markers
```{r , fig.height =3, fig.width =10}
# --- STEP 10F: Treg Activation & Suppressive Phenotype Mapping ---
# 1. Activation Check: Plotting classical markers (Ctla4, Icos, Tigit, Lag3).
# 2. Phenotype Deviation: Checking for 'Fragility' markers (Ifng, Rorc, Il17). 
# High expression here suggests Tregs may be losing their suppressive identity.
DotPlot(subset, group.by = "annot", features = c("Ctla4","Nrp1","Itgae", "Ikzf2","Tnfrsf18","Tnfrsf4","Il2ra","Klrg1","Icos","Ccr8","Tnfsf9","Havcr2","Lag3","Pdcd1","Tigit"))+RotatedAxis()
```


deviation of a suppressive Treg phenotype
```{r , fig.height =3, fig.width =10}
DotPlot(subset, group.by = "annot", features = c("Il17a","Il17b","Il17c","Il17f", "Rorc","Ifng","Il1r1"))+RotatedAxis()
```

```{r , fig.height =4, fig.width =8,warning=FALSE, echo=FALSE}
FeaturePlot(object = subset, c("Il1r2","Bmyc"))
```

## 11. High-Resolution Dimensionality Expansion & Sub-State Sensitivity

### PC=

```{r,warning=FALSE}
# --- STEP 11A: Expanding Dimensionality (PC=25) ---
# We increase the PC threshold to 25 to capture more subtle transcriptional 
# variation that might be masked at lower dimensions. 
# A new UMAP is generated specifically for this PC-depth to compare stability.
dims.use.subset<-25
```



```{r}
subset <- RunUMAP(subset, dims = 1:dims.use.subset, reduction.name = paste0("umapPC",dims.use.subset), reduction.key = paste0("umapPC",dims.use.subset,"_"),verbose =F, n.neighbors = 10) 
```


```{r , fig.height =4, fig.width =8,warning=FALSE, echo=FALSE}
FeaturePlot(object = subset,reduction = paste0("umapPC",dims.use.subset), c("Il1r2","Bmyc"))
```
```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
# --- STEP 11B: Visualizing Cluster Stability ---
# We project our existing 'Annot', 'Detailed', and 'Old' labels onto the new 
# 25-PC UMAP. This allows us to see if the clusters (like Bmyc-hi vs Ccr8-hi) 
# become more distinct or remain stable when more data variance is included.
DimPlot(object = subset, group.by = "annot.old",reduction = paste0("umapPC",dims.use.subset), label=T, repel=T, cols=cols.use)
```

```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, group.by = "annot",reduction = paste0("umapPC",dims.use.subset), label=T, repel=T, cols=cols.use)
```
```{r , fig.height =5, fig.width =7,warning=FALSE, echo=FALSE}
DimPlot(object = subset, group.by = "annot.detailed",reduction = paste0("umapPC",dims.use.subset), label=T, repel=T, cols=cols.use)
```


# Re-evaluating the primary activated Treg 'flavors' at high resolution:
# 1. Ccr8-hi / Il10-hi: The highly suppressive, terminally activated state.
# 2. Bmyc-hi / Cd70-hi: The metabolically distinct and potentially 'fragile' state.
# Using 'umapPC25' ensures we are seeing the most refined spatial separation possible.

Il1r2-hi Ccr8-hi
```{r , fig.height = 8, fig.width = 12,warning=FALSE}
FeaturePlot(object = subset,  features = c("Ccr8",  "Ccl4" ,"Il10" , "Bcl2a1d" , "Vsir" ,"Irf8", "Glipr1"),reduction = paste0("umapPC",dims.use.subset), ncol=3)
```

Il1r2-hi Bmyc-hi
```{r , fig.height = 8, fig.width = 12,warning=FALSE}
FeaturePlot(object = subset,reduction = paste0("umapPC",dims.use.subset),  features = c( "Bmyc" ,"Cd83"  ,  "Nrip1" , "Chchd10" , "Gda","Hspa9"  ,  "Dgat2" ,"Cd70", "Gadd45g"   ), ncol=3)
```


## 12. Fine-Grained Clustering & Stability Analysis (PC=25)

#### Clustering
```{r ,warning=FALSE}
subset <- FindNeighbors(subset, dims = 1:dims.use.subset, verbose=F, k.param=10)
for ( i in seq(0,2, 0.25))
subset <- FindClusters(subset, resolution = i, verbose=F, algorithm = 4)
# algorithm= 4 is Leiden algorithm - often performs better
```



#### Plot of a clustering tree showing the relationship between clusterings at different resolutions. (using the clustree package)
```{r , fig.height = 8, fig.width = 8,warning=FALSE}
clustree(subset, prefix = "RNA_snn_res.")+
   ggtitle("Clustering tree")
```
```{r , fig.height = 8, fig.width = 8,warning=FALSE}
clustree(subset, prefix = "RNA_snn_res.",node_colour = "Il1r2", node_colour_aggr = "median")+
   ggtitle("Clustering tree: Il1r2")
```

```{r , fig.height = 6, fig.width =8,warning=FALSE}
plot<-list()
for ( res in c( 0.75,1,1.25,1.5))
plot[[as.character(res)]]<-DimPlot(subset, pt.size = 1,label=T,repel=T, group.by = paste0("RNA_snn_res.",res), cols=cols.use,reduction = paste0("umapPC",dims.use.subset)) +
   ggtitle(paste("res=",res))
plot_grid(plotlist=plot)
```


```{r ,warning=FALSE}
# --- STEP 12B: Resolution Selection and UMAP Projection ---
# Based on stability, resolution 1.5 is chosen (yielding 9 clusters).
# These clusters are projected onto the High-PC UMAP (umapPC25) to ensure 
# spatial separation matches mathematical grouping.
res=1.5
Idents(subset)=paste0("RNA_snn_res.",res)
Idents(subset)=factor(Idents(subset), levels=1:9)
```
```{r , fig.height = 3, fig.width =4,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use,reduction = paste0("umapPC",dims.use.subset)) 
```
```{r , fig.height =4, fig.width =8,warning=FALSE, echo=FALSE}
FeaturePlot(object = subset,reduction = paste0("umapPC",dims.use.subset), c("Il1r2","Bmyc"))
```

```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"))+RotatedAxis()
```


```{r}
DEgenes=list()
for (i in levels(Idents(subset))){
  DEgenes[[i]]<-FindMarkers(subset, ident.1 = i,min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000)
  DEgenes[[i]]$cluster=i
  DEgenes[[i]]$pseudocount=0.01
  DEgenes[[i]]$max.cells.per.ident=1000
  DEgenes[[i]]$gene=row.names(DEgenes[[i]])
}
saveRDS(DEgenes, paste0( path.output,"DEgenes_PC",dims.use.subset,".res",res,"_", subset_name,"_",dataset.name,".rds"))
```


```{r}
res=1.5
dims.use.subset=25
DEgenes <-readRDS( paste0( path.output,"DEgenes_PC",dims.use.subset,".res",res,"_", subset_name,"_",dataset.name,".rds"))
```

hypoxia
```{r , fig.height =3, fig.width =13}
# --- STEP 12C: Automated & Hypoxia-Targeted Marker Discovery ---
# 1. Hypoxia Check: A dedicated DotPlot for glycolytic and oxygen-response genes
#    (Vegfa, Eno1, Ldha) determines the metabolic state of each cluster.
# 2. Top-DE Discovery: We extract the top 7 biomarkers for every cluster to 
#    identify unique 'fingerprint' genes.
DotPlot(subset, features = c("Hilpda", "Vegfa", "Bnip3", "Hmox1", "Fth1", "Aldoa", "Mif", "Ftl1", "Arg1", "Ldha", "Adam8", "Eno1", "Srgn", "Ndrg1", "Mt1", "Fam162a", "Slc7a2", "Slc2a1", "Mgarp", "Lgals3","Spp1" ))+RotatedAxis()+theme(axis.title = element_blank())
```
#### Check the expression of the top DE genes using a DotPlot
```{r}
features.use=unlist(lapply(DEgenes, function(x) { head(x[x$avg_logFC>0,]$gene,7)}))
length(features.use)
names(features.use)=NULL
features.use=features.use[!duplicated(features.use)]
```

```{r , fig.height =4, fig.width =14}
DotPlot(subset, features = features.use)+RotatedAxis()
```
```{r , fig.height = 3, fig.width =4,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use,reduction = paste0("umapPC",dims.use.subset)) 
```


```{r , fig.height =4, fig.width =14}
DotPlot(subset, features = c( "Ifngr1","Icos", "Il2ra","Foxp3",  "Il7r","Btg1","Ccr2","Gzmb","Ccl5","Fus","Ctla4","perc.ribo","Rpl17","Fau",  "Hilpda" , "Ddit3","Hspa1b","Hspa1a"   , "Eno1" ,"Bnip3" ,"Cst7" ,"Bcl2a1b","Nkg7" ,"Csf1","Rel"  ,"Ly6a","Ccr8","Il1r2","Bmyc" ,"Mif","Ddt","Slc39a7","Dnajb14","Zfp593","Nhp2","Actr5","Tff1","Myo18a","Scnm1", "Chchd10","Cd83" ,  "Nrip1" ,"Abcb10","Ccrl2","Slfn8","Ifit1","Icam1",  "Ifi203","Ifit3","Isg15" ,"Shisa5" ,"Ran","Mki67","Stmn1","doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```


```{r ,warning=FALSE}
# --- STEP 12D: Final Cluster Identity Mapping ---
# Numbers are mapped to detailed biological states. Note the distinction 
# between 'Treg Il1r2-hi IFN' (Activated + Fragile) and 'Treg Il1r2-hi Bmyc-hi'.
# CellSelector is used to manually 'clean' the proliferating cluster based on Mki67.
Idents(subset)=plyr::mapvalues(Idents(subset), from=1:9, 
    to=c("Treg Ifngr1","Treg Il1r2-hi IFN","Treg Il1r2-hi ribo-hi","Treg Il1r2-hi prolif","Treg Ifngr1 Ccr2",
         "Treg Il1r2-hi Bmyc-hi","Treg Csf1","Treg IFN","Treg Csf1"))
```





```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use,reduction = paste0("umapPC",dims.use.subset)) 
```

```{r , fig.height =4, fig.width =6,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use, group.by = "annot.old",reduction = paste0("umapPC",dims.use.subset)) 
```


```{r , fig.height =8, fig.width =10,warning=FALSE}
FeaturePlot(subset,c("Mki67","Stmn1","Mcm3","Il1r2"),reduction = paste0("umapPC",dims.use.subset)) 
```

select.cells=CellSelector(FeaturePlot(subset,c("Mki67"),reduction = paste0("umapPC",dims.use.subset)))
Idents(subset, cells=select.cells)="Treg Il1r2-hi prolif"
subset$annot.detailed=Idents(subset)


```{r}
new.order=c ( "Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2","Treg IFN" ,"Treg Il1r2-hi IFN", "Treg Il1r2-hi ribo-hi", "Treg Il1r2-hi Bmyc-hi" ,
        "Treg Il1r2-hi prolif"    )
new.order[!new.order %in% levels( Idents(subset))]
levels( Idents(subset))[!levels( Idents(subset)) %in%new.order]
```



```{r}
Idents(subset)=factor( Idents(subset), levels=new.order)
subset$annot.detailed=Idents(subset)
```



```{r}
# --- STEP 12E: Lineage Validation (CD4 T vs. Treg) ---
# A final high-level comparison between general CD4 T cells and Tregs.
# This validates that our subsetting was accurate and identifies the 
# global 'Treg Signature' (Foxp3, Il2ra, etc.) across the tumor.
DEG=FindMarkers(seur_full, ident.1 = "CD4 T", ident.2 = "Treg")
DEG$gene=rownames(DEG)
```

```{r , fig.height =4, fig.width =12}
VolcanoPlot(DEG)
```



```{r}
selected.genes=DEG[DEG$p_val<0.1 &  DEG$avg_logFC>0.5,"gene"]
length(selected.genes)
```
```{r , fig.height =4, fig.width =12}
DotPlot(seur_full, group.by = "annot", features = unique(c(selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = unique(c(selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```

## 13. Cross-Dataset Signature Projection & Bmyc-hi Sub-cluster Validation

```{r}
# --- STEP 13A: Projecting External MC38 Signatures ---
# We use gene lists from the MC38 Treg reference (Bmyc-hi, Ccr8-hi, and Ctla4 signatures)
# to see how well they align with our current clusters. This serves as a 
# 'biological ground truth' for our detailed annotations.
i="Treg.Il1r2.hi.Bmyc.hi"
selected.genes= DEgenes_MC38_Tregs [[i]][ DEgenes_MC38_Tregs[[i]]$p_val<0.01 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.35,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```



```{r}
# --- STEP 13B: Validating Specific Signature Modules ---
# We repeat the projection for the 'Ccr8-hi' and 'Ctla4' reference signatures.
# This confirms that our activated clusters are not monolithic but represent 
# distinct functional programs found in other tumor models.
i= "Treg.Il1r2.hi.Ccr8.hi"
selected.genes= DEgenes_MC38_Tregs[[i]][ DEgenes_MC38_Tregs[[i]]$p_val<0.01 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.3,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r}
i="Treg.Ctla4"  
selected.genes= DEgenes_MC38_Tregs[[i]][ DEgenes_MC38_Tregs[[i]]$p_val_adj<1e-3 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.5,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c(selected.genes))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(subset,label=T,repel=T,cols= cols.use,reduction = paste0("umapPC",dims.use.subset)) 
```
```{r , fig.height =4, fig.width =14}
DotPlot(subset, features = c( "Foxp3","Il2ra","Ctla4","Cst7" ,"Bcl2a1b","Nkg7" ,"Csf1","Rel"  , "Btg1","Ifngr1","Icos",  "Il7r","Ccr2","Gzmb","Ccl5","Fus","Ifit3","Slfn8","Ifit1",  "Ifi203","Isg15" ,"Gbp7" ,"Ly6a","Il1r2","perc.ribo","Rpl17","Fau",  "Hilpda" ,"Chchd10","Bmyc" , "Ddt","Slc39a7","Dnajb14","Tff1","Ripk3","Hmgn5","Scnm1", "Zfp593","Nhp2","Mif","Ran","Mki67","Stmn1","Ccr8","Eno1" ,"Bnip3" ,"doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```


```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"))+RotatedAxis()
```


```{r}
# --- STEP 13C: Broad Marker Profiling Across Annotations ---
# Visualizing the dataset at the 'Detailed' and 'General' levels. This validates 
# that collapsing clusters into 'annot.general' maintains the core distinction 
# between resting Tregs and the highly activated Il1r2-hi/Bmyc-hi subsets.
subset$annot.general=plyr::mapvalues(Idents(subset), from=c( "Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2", "Treg Il1r2-hi ribo-hi" ), to=c("Treg","Treg","Treg", "Treg Il1r2-hi" ))
```



```{r , fig.height =4, fig.width =6,warning=FALSE}
DimPlot(subset, group.by = "annot.general",label=T,repel=T,cols= cols.use,reduction = paste0("umapPC",dims.use.subset)) 
```
```{r , fig.height =4, fig.width =12}
DotPlot(subset, group.by = "annot.general", features = c( "Foxp3","Il2ra","Ctla4","Ifit3","Slfn8","Ifit1",  "Ifi203","Isg15" ,"Gbp7" ,"Ly6a","Il1r2","perc.ribo","Rpl17",  "Hilpda" ,"Bnip3","Chchd10","Bmyc" ,"Tff1","Ripk3","Hmgn5","Scnm1", "Nhp2","Mif","Ran","Mki67","Stmn1","Ccr8","Eno1"  ,"doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```


```{r , fig.height =3, fig.width =5}
DotPlot(subset, group.by = "annot.general", features = c( "Il1r2"))+RotatedAxis()
```


```{r,warning=FALSE, echo=FALSE}
clusters=levels(subset$annot.detailed)
DEgenes_list<-list()
for ( i in clusters){
DEgenes_list[[i]]<-  FindMarkers(subset, group.by="annot.detailed",ident.1=i,min.cells.group=1, max.cells.per.ident = 1000, pseudocount.use = 0.1)
DEgenes_list[[i]]$gene=rownames(DEgenes_list[[i]])
DEgenes_list[[i]]$cluster.1=i
DEgenes_list[[i]]$pseudocount.use=0.1
DEgenes_list[[i]]$max.cells.per.ident=1000
}
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```

```{r,warning=FALSE, echo=FALSE}
DEgenes_list=readRDS( paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.general.rds"))
```
```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"))+RotatedAxis()
```
```{r}
# --- STEP 13D: Identifying the Bmyc-hi Signature ---
# A targeted comparison between the Bmyc-hi cluster and all 'Il1r2-low' 
# (resting/IFN) cells. The resulting Volcano Plot and DotPlot reveal the 
# novel markers (e.g., Tff1, Ripk3, Scnm1) defining this "fragile" state.
i="Il1r2.hi.Bmyc.vs.Il1r2.lo"
  DEgenes_list[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi Bmyc-hi", ident.2 = c("Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2","Treg IFN"), group.by="annot.detailed",min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F, logfc.threshold = 0, min.pct = 0)
  DEgenes_list[[i]]$cluster=i
  DEgenes_list[[i]]$pseudocount=0.01
  DEgenes_list[[i]]$max.cells.per.ident=1000
  DEgenes_list[[i]]$gene=row.names(DEgenes_list[[i]])
  saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```


```{r , fig.height =5, fig.width =12}
i="Il1r2.hi.Bmyc.vs.Il1r2.lo"
VolcanoPlot(  DEgenes_list[[i]], p_valmax_col = 0.5)
```

```{r}
selected.genes=rownames( DEgenes_list[[i]][ DEgenes_list[[i]]$p_val<1e-4 &  DEgenes_list[[i]]$avg_logFC>1,])
length(selected.genes)
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = unique(c( "Il1r2",selected.genes[1:40])))+RotatedAxis()
```
```{r , fig.height =3, fig.width =14}
DotPlot(subset, features =selected.genes[41:93])+RotatedAxis()
```



```{r , fig.height =3, fig.width =12}
# --- STEP 13E: Cross-Reference Mapping (Current vs. MC38) ---
# We take the genes discovered in our Bmyc-hi cluster and project them onto 
# the MC38_Tregs reference dataset. This "reciprocal validation" proves 
# these markers are conserved across different experimental cohorts.
DotPlot(MC38_Tregs, features = unique(c("Il1r2",selected.genes[1:40])))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r , fig.height =3, fig.width =7}
DotPlot(MC38_Tregs, features =c("Il1r2", "Polr3c" ,   "Tff1","Myadm","Mllt11", "Dars","Ripk3","Hmgn5","Scnm1"   ) )+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =8, fig.width =12}
FeaturePlot(MC38_Tregs, features = c("Il1r2", "Polr3c" ,   "Tff1","Myadm","Mllt11", "Dars","Ripk3","Hmgn5","Scnm1"   ), reduction ="umapHarmonyTheta1PC25")+RotatedAxis()+theme(axis.title = element_blank())
```


```{r , fig.height =4, fig.width = 6}
DimPlot(MC38_Tregs,label=T,repel=T, cols=cols.use,reduction ="umapHarmonyTheta1PC25")
```

```{r , fig.height =8, fig.width =12}
FeaturePlot(subset, features = c("Il1r2", "Polr3c" ,   "Tff1","Myadm","Mllt11", "Dars","Ripk3","Hmgn5","Scnm1"  ) ,reduction = paste0("umapPC",dims.use.subset))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =4, fig.width = 4}
# --- STEP 13F: Highlighted Spatial Localization ---
# Final visualization isolating the 'Treg Il1r2-hi Bmyc-hi' cluster on the 
# PC25 UMAP. This identifies the specific transcriptional "territory" of 
# these cells within the total Treg population.
DimPlot(subset,cells.highlight = WhichCells(subset, idents="Treg Il1r2-hi Bmyc-hi"), sizes.highlight = 0.8,reduction = paste0("umapPC",dims.use.subset))+NoLegend()
```



Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```


## 14. Characterization of Ribosomal and Interferon-Activated Subsets

```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```

```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
# --- STEP 14A: Pathway Projection for Bmyc-hi Subsets ---
# Projecting the top enriched GO terms from Metascape onto the Bmyc-hi vs. Il1r2-lo 
# volcano plots. This connects differential gene expression to high-level biological 
# processes such as 'ribonucleoprotein complex biogenesis' or 'cellular respiration'.
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=2
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```


```{r}
# --- STEP 14B: Defining the Ribosomal-High Treg Program ---
# Performing a targeted comparison between the 'Treg Il1r2-hi ribo-hi' cluster 
# and the resting 'Il1r2-lo' populations. This identifies a unique signature 
# of high protein synthesis (Rps7, Rpl32) and metabolic stress (Hilpda, Ddit4).
i="Il1r2.hi.ribo.vs.Il1r2.lo"
  DEgenes_list[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi ribo-hi", ident.2 = c("Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2","Treg IFN"), group.by="annot.detailed",min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes_list[[i]]$cluster=i
  DEgenes_list[[i]]$pseudocount=0.01
  DEgenes_list[[i]]$max.cells.per.ident=1000
  DEgenes_list[[i]]$gene=row.names(DEgenes_list[[i]])
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```


```{r , fig.height =5, fig.width =12}
i="Il1r2.hi.ribo.vs.Il1r2.lo"
VolcanoPlot(  DEgenes_list[[i]], p_valmax_col = 0.5, logFCmin_col = 0.5)
```

```{r}
selected.genes=rownames( DEgenes_list[[i]][ DEgenes_list[[i]]$p_val<1e-3 &  DEgenes_list[[i]]$avg_logFC>0.5,])
length(selected.genes)
```


```{r , fig.height =3, fig.width =14}
DotPlot(subset, features = unique(c( "Il1r2",selected.genes)))+RotatedAxis()
```

```{r , fig.height =8, fig.width =12}
# --- STEP 14C: Cross-Dataset Validation of the Ribo-hi State ---
# Mapping the newly identified ribo-hi markers (Vim, Uba52, Mrpl32) across 
# both the current subset and the MC38_Tregs reference. This validates 
# that this high-translation effector state is a conserved Treg phenotype.
FeaturePlot(subset, features = c("Rps7","Rpl32","Vim","Uba52","Mrpl32", "Mrpl11","Tff1" ,   "Hilpda", "Emd","Ddit4","Ppil3","Ero1l"  ) ,reduction = paste0("umapPC",dims.use.subset))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =4, fig.width = 4}
# --- STEP 14D: Isolating the 'Ribo-hi' Territory ---
# Using cells.highlight to pinpoint exactly where these high-translation cells 
# reside on the high-resolution UMAP, confirming they occupy a distinct 
# spatial niche from other activated Tregs.
DimPlot(subset,cells.highlight = WhichCells(subset, idents="Treg Il1r2-hi ribo-hi" ), sizes.highlight = 0.8,reduction = paste0("umapPC",dims.use.subset))+NoLegend()
```

```{r , fig.height =8, fig.width =12}
FeaturePlot(MC38_Tregs, features = c("Rps7","Rpl32","Vim","Uba52","Mrpl32", "Mrpl11","Tff1" ,   "Hilpda", "Emd","Ddit4","Ppil3","Ero1l"  ), reduction ="umapHarmonyTheta1PC25")+RotatedAxis()+theme(axis.title = element_blank())
```


```{r , fig.height =4, fig.width = 6}
DimPlot(MC38_Tregs,label=T,repel=T, cols=cols.use,reduction ="umapHarmonyTheta1PC25")
```


Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```


```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```


```{r}
# --- STEP 14E: Defining the Activated-Interferon (IFN) Program ---
# A focused comparison between 'Treg Il1r2-hi IFN' and the resting populations. 
# While both 'Treg IFN' and 'Treg Il1r2-hi IFN' clusters respond to interferon, 
# this analysis extracts the specific genes that signal high activation 
# (Il1r2) combined with an interferon-induced inflammatory state.
i="Il1r2.hi.IFN.vs.Il1r2.lo"
  DEgenes_list[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi IFN", ident.2 = c("Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2","Treg IFN"), group.by="annot.detailed",min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes_list[[i]]$cluster=i
  DEgenes_list[[i]]$pseudocount=0.01
  DEgenes_list[[i]]$max.cells.per.ident=1000
  DEgenes_list[[i]]$gene=row.names(DEgenes_list[[i]])
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```



```{r , fig.height =5, fig.width =12}
i="Il1r2.hi.IFN.vs.Il1r2.lo"
VolcanoPlot(  DEgenes_list[[i]], p_valmax_col = 0.5, logFCmin_col = 0.5)
```

```{r}
selected.genes=rownames( DEgenes_list[[i]][ DEgenes_list[[i]]$p_val<1e-3 &  DEgenes_list[[i]]$avg_logFC>0.5,])
length(selected.genes)
```

```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c( "Il1r2",selected.genes))+RotatedAxis()
```


Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```

## 15. Refined IFN Mapping and Global Activated Treg Signatures

```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"/metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```


```{r}
# --- STEP 15A: Specific Interferon Activation vs. Resting Lineages ---
# To find the 'activation-specific' interferon signal, we compare the 'Treg Il1r2-hi IFN' 
# cluster against resting Tregs (Csf1, Ifngr1, Ccr2) while excluding the baseline 
# 'Treg IFN' cluster. This isolates genes that are induced by interferon 
# specifically within a highly activated (Il1r2+) context.
i="Il1r2.hi.IFN.vs.Il1r2.lo.exclIFN"
  DEgenes_list[[i]]<-FindMarkers(subset, ident.1 = "Treg Il1r2-hi IFN", ident.2 = c("Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2"), group.by="annot.detailed",min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F)
  DEgenes_list[[i]]$cluster=i
  DEgenes_list[[i]]$pseudocount=0.01
  DEgenes_list[[i]]$max.cells.per.ident=1000
  DEgenes_list[[i]]$gene=row.names(DEgenes_list[[i]])
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```



```{r , fig.height =5, fig.width =12}
i="Il1r2.hi.IFN.vs.Il1r2.lo.exclIFN"
VolcanoPlot(  DEgenes_list[[i]], p_valmax_col = 1, logFCmin_col = 0.5)
```

```{r}
selected.genes=rownames( DEgenes_list[[i]][ DEgenes_list[[i]]$p_val<1e-2 &  DEgenes_list[[i]]$avg_logFC>0.5,])
length(selected.genes)
```

```{r , fig.height =3, fig.width =12}
DotPlot(subset, features = c( "Il1r2",selected.genes[1:50]))+RotatedAxis()
```


Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```


```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
# --- STEP 15B: GO Enrichment Projection (IFN Refined) ---
# Projecting the Metascape-derived 'Summary' pathways back onto the volcano plots. 
# This confirms that the genes driving the Il1r2-hi IFN state are indeed 
# enriched for 'Response to Type I Interferon' and 'Antiviral Response' programs.
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
gset.symbols
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols, dot.size = 2, max.overlaps.use=20)+ggtitle(paste0( i, ": ", GOterm))
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=2
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```


```{r}
# --- STEP 15C: Defining the Global 'Il1r2-hi' Activated Program ---
# We group all three specialized activated clusters (IFN, Bmyc-hi, ribo-hi) 
# to compare them against all resting/baseline clusters (Il1r2-lo). 
# This yields the definitive 'Master Signature' of highly suppressive, 
# tumor-infiltrating Tregs.
i="Il1r2.hi.all.vs.Il1r2.lo"
  DEgenes_list[[i]]<-FindMarkers(subset, ident.1 =c( "Treg Il1r2-hi IFN","Treg Il1r2-hi Bmyc-hi","Treg Il1r2-hi ribo-hi"), ident.2 = c("Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2","Treg IFN"), group.by="annot.detailed",min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F, logfc.threshold = 0, min.pct = 0)
  DEgenes_list[[i]]$cluster=i
  DEgenes_list[[i]]$pseudocount=0.01
  DEgenes_list[[i]]$max.cells.per.ident=1000
  DEgenes_list[[i]]$gene=row.names(DEgenes_list[[i]])
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```


```{r , fig.height =5, fig.width =12}
# --- STEP 15D: Visualization of the Core Activation Signature ---
# Using VolcanoPlots and DotPlots to visualize the high-confidence genes (p < 1e-3, logFC > 0.5) 
# that define the global activated state. This reveals shared markers like Il1r2, 
# Ccr8, and various metabolic or stress-related genes across all activated sub-states.
i="Il1r2.hi.all.vs.Il1r2.lo"
VolcanoPlot(  DEgenes_list[[i]], p_valmax_col = 0.3, logFCmin_col = 0.3, logFCmin_neg_col = -0.3)
```

```{r}
selected.genes=rownames( DEgenes_list[[i]][ DEgenes_list[[i]]$p_val<1e-3 &  DEgenes_list[[i]]$avg_logFC>0.5,])
length(selected.genes)
```

```{r , fig.height =3, fig.width =15}
DotPlot(subset, features =unique( c( "Il1r2",selected.genes)))+RotatedAxis()
```


Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```

## 16. Multi-Pathway Enrichment and Proliferative Signature Analysis

```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"/metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
# --- STEP 16A: Multi-Index GO Pathway Projection ---
# We systematically project the top four enriched biological pathways (Index 1-4) 
# from the Metascape 'Summary' list onto the Volcano plots. This provides 
# statistical proof that the differentially expressed genes are not random, 
# but are coordinated across specific biological functions (e.g., cell cycle control, 
# metabolism, or immune signaling).
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```

```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=2
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```

```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=3
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```

```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
index=4
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```


```{r}
# --- STEP 16B: Isolating the Proliferative Activation State ---
# We perform a targeted differential expression analysis between the 
# 'Treg Il1r2-hi prolif' cluster and the resting 'Il1r2-lo' lineages. 
# This isolates the unique genetic program required for Tregs to transition 
# from a quiescent state to an actively dividing, highly activated state.
i="Il1r2.ih.prolifl.vs.Il1r2.lo"
  DEgenes_list[[i]]<-FindMarkers(subset, ident.1 =c( "Treg Il1r2-hi prolif"), ident.2 = c("Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2","Treg IFN"), group.by="annot.detailed",min.cells.group=2,pseudocount.use = 0.01, max.cells.per.ident = 1000,verbose = F, logfc.threshold = 0, min.pct = 0)
  DEgenes_list[[i]]$cluster=i
  DEgenes_list[[i]]$pseudocount=0.01
  DEgenes_list[[i]]$max.cells.per.ident=1000
  DEgenes_list[[i]]$gene=row.names(DEgenes_list[[i]])
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```


```{r , fig.height =5, fig.width =12}
i="Il1r2.ih.prolifl.vs.Il1r2.lo"
VolcanoPlot(  DEgenes_list[[i]], p_valmax_col = 0.05, logFCmin_col = 1, logFCmin_neg_col = -0.3)
```

```{r}
# --- STEP 16C: High-Confidence Marker Selection (Proliferation) ---
# Filtering for the most significant genes (p < 1e-8, logFC > 1) associated 
# with the proliferative subset. Using DotPlots to confirm these markers 
# (e.g., Mki67, Stmn1, Top2a) are exclusive to this cluster within the 
# activated Treg landscape.
selected.genes=rownames( DEgenes_list[[i]][ DEgenes_list[[i]]$p_val<1e-8 &  DEgenes_list[[i]]$avg_logFC>1,])
length(selected.genes)
```

```{r , fig.height =4, fig.width =15}
DotPlot(subset, features =unique( c( "Il1r2",selected.genes[1:50])))+RotatedAxis()
```


Export gene names as a csv file for metascape
```{r}
write.table(selected.genes, file=paste0(path.output,"DEG_",make.names(i),"_",subset_name,"_",dataset.name,".csv"), row.names = F, col.names = F, sep=",")
```

## 17. Final Biological Validation and Multi-Level Data Export

```{r}
GO.metascape=readxl::read_xlsx( paste0(path.output,"/metascape_DEG_",make.names(i),"_",subset_name,"_",dataset.name,".xlsx"),sheet=2 )
#Select only the main pathways
GO.metascape$group=sapply(strsplit(GO.metascape$GroupID, "_"), "[[",2)
GO.metascape.summary=GO.metascape[GO.metascape$group=="Summary",]
```


```{r , fig.height =4, fig.width = 10 ,warning=FALSE}
# --- STEP 17A: Final GO Pathway Projection ---
# Mapping the primary enriched biological pathway from Metascape onto the current 
# Volcano Plot. This step provides the final visual confirmation that the 
# differentially expressed genes in the chosen cluster represent a coherent 
# functional program (e.g., T cell activation or metabolic reprogramming).
index=1
gset.symbols=GO.metascape.summary[index,]$Symbols
gset.symbols=strsplit(gset.symbols, ",")[[1]]
GOterm=GO.metascape.summary[index,"Description"]
VolcanoPlot(DEgenes_list[[i]],subset_col_gene=gset.symbols)+ggtitle(paste0( i, ": ", GOterm))
```

```{r,warning=FALSE, echo=FALSE}
# --- STEP 17B: Archiving Detailed Cluster DEGs ---
# We save the high-resolution 'annot.detailed' markers into a permanent RDS file 
# and a formatted Excel workbook. This serves as the primary data resource for 
# describing specific Treg sub-states (Bmyc-hi, Ccr2-hi, etc.) in the manuscript.
DEgenes_list=readRDS( paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.rds"))
```


```{r,warning=FALSE, echo=FALSE}
names(DEgenes_list)=make.names(names(DEgenes_list))
### Save the differentially expressed genes as an excel table
writexl::write_xlsx(DEgenes_list,paste0(path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.detailed.xlsx"))
```


```{r,warning=FALSE, echo=FALSE}
# --- STEP 17C: Generating General Population Markers ---
# To provide a 'higher-level' view of Treg activation, we perform differential 
# expression across the broad 'annot.general' groups (e.g., Treg vs. Treg Il1r2-hi). 
# This isolates the fundamental markers of activation that are shared across 
# all specialized sub-clusters.
clusters=levels(subset$annot.general)
DEgenes_list<-list()
for ( i in clusters){
DEgenes_list[[i]]<-  FindMarkers(subset, group.by="annot.general",ident.1=i,min.cells.group=1, max.cells.per.ident = 1000, pseudocount.use = 0.1)
DEgenes_list[[i]]$gene=rownames(DEgenes_list[[i]])
DEgenes_list[[i]]$cluster.1=i
DEgenes_list[[i]]$pseudocount.use=0.1
DEgenes_list[[i]]$max.cells.per.ident=1000
}
saveRDS(DEgenes_list, paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.general.rds"))
```



```{r , fig.height =3.5, fig.width =5}
# --- STEP 17D: Final Activation Gradient Check ---
# A final DotPlot focuses strictly on the expression of Il1r2 across the 
# generalized groups. This ensures that our "General" labels accurately 
# represent the activation rheostat identified throughout the analysis.
DotPlot(subset, group.by = "annot.general", features = c( "Il1r2"))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r,warning=FALSE, echo=FALSE}
DEgenes_list=readRDS( paste0( path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.general.rds"))
```
```{r,warning=FALSE, echo=FALSE}
names(DEgenes_list)=make.names(names(DEgenes_list))
### Save the differentially expressed genes as an excel table
writexl::write_xlsx(DEgenes_list,paste0(path.output,"DEgenes_", subset_name,"_",dataset.name,".annot.general.xlsx"))
```




## 18. Global Integration of Detailed Treg Annotations


```{r ,warning=FALSE}
# --- STEP 18A: Global Metadata Synchronization ---
# We transfer the high-resolution 'annot.detailed' labels from the Treg subset 
# back into the full Seurat object (seur_full). This ensures that the 
# discoveries made during the subsetting phase are reflected in the 
# master clinical dataset.
Idents(seur_full)="annot"
for ( i in unique(Idents(subset)))
  Idents(seur_full, WhichCells(subset, idents=i))=i
seur_full$annot.detailed=Idents(seur_full)
```


```{r , fig.height = 6, fig.width =10,warning=FALSE}
# --- STEP 18B: Final Multi-Populated UMAP Projection ---
# Generating the master UMAP for the entire experiment. This visualization 
# places the detailed Treg populations within the context of the total 
# tumor microenvironment. It provides the final spatial evidence for 
# how activated Tregs are distributed relative to other immune subsets.
DimPlot(seur_full, group.by =  "annot.detailed",label=T,repel=T, cols=cols.use) 
```


## 19. T/NK Cell Subset Integration and Refined Preprocessing

### subset T/NK cells

```{r}
# --- STEP 19A: Lymphoid Compartment Subsetting ---
# We isolate all T cell (CD4, CD8, Treg) and NK cell populations from the master 
# dataset. This 'lymphocyte-only' view removes the high variance introduced by 
# myeloid or stromal cells, allowing for a more sensitive UMAP projection of 
# lymphoid states.
seur=seur_full[,WhichCells(seur_full, idents=c("Treg Il1r2-hi prolif" , "Treg IFN" ,  "Treg Il1r2-hi ribo-hi" ,"Treg Il1r2-hi IFN",   "Treg Csf1" ,         "Treg Ifngr1" , "Treg Il1r2-hi Bmyc-hi", "Treg Ifngr1 Ccr2",    "Treg"  ,    "CD4T"    ,    "CD8T"  ,     "T prolif" , "NK"       ))]
```


```{r,warning=FALSE}
dim(seur)
```


 Standard preprocessing and clustering of the subsetted dataset
```{r,warning=FALSE}
# --- STEP 19B: Specialized Lymphoid Normalization & PCA ---
# Standard preprocessing is re-applied to the subset. The Elbow Plot and 
# DimHeatmaps are used to determine the optimal number of Principal Components 
# (PCs). For this subset, 21 PCs are selected to capture the relevant 
# variance between T and NK cell lineages.
seur <-  NormalizeData(seur,verbose = F) %>% FindVariableFeatures(verbose = F) %>% ScaleData(verbose = F) %>% RunPCA( features = VariableFeatures(seur) ,verbose = F)
```
```{r , fig.height = 4, fig.width = 6,warning=FALSE}
ElbowPlot(object = seur,ndims =50)
```
```{r , fig.height = 10, fig.width = 12}
DimHeatmap(seur, dims = 1:21, balanced = TRUE)
```
```{r , fig.height = 10, fig.width = 12}
DimHeatmap(seur, dims = 22:39, balanced = TRUE)
```
```{r,warning=FALSE}
# --- STEP 19C: UMAP Projection & General Annotation Mapping ---
# The UMAP is generated using the refined 21-PC space. We then map our 
# 'annot.general' labels onto this new space to verify that the Treg 
# activation states remain distinct within the broader lymphocyte context.
dims.use<-21
```



```{r}
seur <- RunUMAP(seur, dims = 1:dims.use,verbose =F ) 
```

```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(seur,label=T,repel=T,cols= cols.use) 
```



```{r}
seur$annot.general=plyr::mapvalues(Idents(seur), from=c( "Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2", "Treg Il1r2-hi ribo-hi" ), to=c("Treg","Treg","Treg", "Treg Il1r2-hi" ))
```


```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(seur,label=T,repel=T,cols= cols.use, group.by ="annot.general" ) 
```
select.cells=CellSelector(DimPlot(seur,cols= cols.use, group.by ="annot.general" ) )
Idents(seur, cells=select.cells)="NK2"

```{r }
# --- STEP 19D: Discovery of Novel NK Subsets (NK2) ---
# Using 'CellSelector' to manually identify a distinct island on the UMAP 
# (designated as 'NK2'). We then perform Differential Expression (DE) between 
# this new group and the standard NK cells to identify unique markers or 
# potential doublet contamination.
DEG=FindMarkers(seur, ident.1 = "NK2", ident.2 = "NK")
```
```{r , fig.height =4, fig.width =12}
VolcanoPlot(DEG)
```
```{r , fig.height = 4, fig.width =13}
# --- STEP 19E: Multi-Lineage Marker Profiling ---
# A comprehensive DotPlot is used to 'fingerprint' every cluster in the T/NK subset. 
# This includes markers for Tregs (Foxp3), CD8s (Cd8a), NK cells (Ncr1), and 
# potential contaminants (Lyz2 for Myeloid, Cd79a for B cells).
DotPlot(seur,features=c("Cd3e","Cd3g","Foxp3","Il2ra","Tnfrsf9","Ikzf2","Izumo1r","Mki67","Stmn1","Cd4","Cd40lg","S1pr1","Cd8a","Cd8b1","Xcl1","Klra7","Klrb1c","Ncr1","Gzma","Cd79a","Cd79b","Cd19","Ms4a1","Csf3r","Retnlg","S100a8","Lyz2","Fn1","Chil3","F13a1","Vcan","Tgfbi","Adgre1","C1qa","Ms4a7","Lgmn","H2-Aa","Cd74","H2-DMb1","Cd209a","Napsa","Kmo","Ccr9","Siglech","Pacsin1","Xcr1","Rab7b","Clec9a","Flt3","Cacnb3","Ccl22","Nudt17","doublet.score"  ))+RotatedAxis()+theme(axis.title=element_blank())
```
```{r , fig.height =8, fig.width =12,warning=FALSE}
# --- STEP 19F: Spatial Doublet & Contamination Check ---
# FeaturePlots for Myeloid markers (C1qa, Lyz2) and 'doublet.score' are mapped 
# onto the lymphoid UMAP. This is a critical QC step to ensure that small 
# clusters (like the newly identified NK2) are true biological populations 
# rather than T-cell/Macrophage doublets.
FeaturePlot(seur,c("C1qa","C1qb","Lyz2","Chil3","Vcan","H2-Aa","Cd209a","Csf3r","doublet.score")) 
```

## 20. Subset Refinement and Hierarchical T/NK Cell Mapping

```{r }
# --- STEP 20A: Cleaning the Lymphoid Compartment ---
# Based on previous QC markers, the "NK2" population is removed (invert=T). 
# This ensures that downstream PCA and UMAP projections are driven by true 
# lymphoid biology rather than leftover noise or doublets.
seur=seur[,WhichCells(seur, idents="NK2", invert=T)]
```

```{r,warning=FALSE}
# --- STEP 20B: Re-calculation of the PCA/UMAP Space ---
# A second round of normalization and Scaling is performed on the clean subset. 
# We maintain 21 PCs (as verified by the Elbow Plot and DimHeatmap) to preserve 
# the fine-grained differences between T cell effector states.
seur <-  NormalizeData(seur,verbose = F) %>% FindVariableFeatures(verbose = F) %>% ScaleData(verbose = F) %>% RunPCA( features = VariableFeatures(seur) ,verbose = F)
```
```{r , fig.height = 4, fig.width = 6,warning=FALSE}
ElbowPlot(object = seur,ndims =50)
```
```{r , fig.height = 10, fig.width = 12}
DimHeatmap(seur, dims = 1:21, balanced = TRUE)
```
```{r , fig.height = 10, fig.width = 12}
DimHeatmap(seur, dims = 22:39, balanced = TRUE)
```
```{r,warning=FALSE}
dims.use<-21
```



```{r}
seur <- RunUMAP(seur, dims = 1:dims.use,verbose =F ) 
```

```{r}
new.order=c ("NK",  "CD8T" ,"T prolif" , "CD4T", "Treg","Treg Csf1", "Treg Ifngr1","Treg Ifngr1 Ccr2"  ,"Treg IFN","Treg Il1r2-hi IFN" ,"Treg Il1r2-hi ribo-hi","Treg Il1r2-hi Bmyc-hi","Treg Il1r2-hi prolif"             )
new.order[!new.order %in% levels( Idents(seur))]
levels( Idents(seur))[!levels( Idents(seur)) %in%new.order]
```



```{r}
# --- STEP 20C: Hierarchical Annotation Mapping ---
# We establish a three-tier metadata system to allow for analysis at different 
# levels of granularity:
# 1. annot.detailed: The most specific labels (e.g., Treg Il1r2-hi Bmyc-hi).
# 2. annot: Collapses specific ribosomal and resting Treg clusters.
# 3. annot.general: Groups all activated Il1r2-hi cells together for 
#    broad population comparisons.
Idents(seur)=factor( Idents(seur), levels=new.order)
seur$annot.detailed=Idents(seur)
seur$annot=plyr::mapvalues(Idents(seur), from=c( "Treg Csf1","Treg Ifngr1","Treg Ifngr1 Ccr2", "Treg Il1r2-hi ribo-hi" ), to=c("Treg","Treg","Treg", "Treg Il1r2-hi" ))
seur$annot.general=plyr::mapvalues(seur$annot, from=c( "Treg Il1r2-hi prolif","Treg Il1r2-hi IFN" ,"Treg Il1r2-hi Bmyc-hi"  ), to=c( "Treg prolif","Treg Il1r2-hi" ,"Treg Il1r2-hi"  ))
```


```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(seur,label=T,repel=T,cols= cols.use) 
```
```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(seur,label=T,repel=T,cols= cols.use, group.by ="annot" ) 
```
```{r , fig.height =6, fig.width =7,warning=FALSE}
DimPlot(seur,label=T,repel=T,cols= cols.use, group.by ="annot.general" ) 
```

```{r , fig.height = 4, fig.width =12}
# --- STEP 20D: Global Lineage Verification ---
# DotPlots are used to confirm that every cluster expresses its expected 
# lineage markers: NK (Ncr1, Gzma), CD8 (Cd8a), CD4 (Cd40lg), and Treg (Foxp3). 
# This "Master DotPlot" provides the primary proof of cluster identity.
DotPlot(seur,features=c("Xcl1","Klra7","Klrb1c","Ncr1","Gzma" ,"Cd3e","Cd3g","Cd8a","Cd8b1", "Cd4","Cd40lg","S1pr1","Foxp3","Il2ra","Ctla4","Cst7" ,"Bcl2a1b","Nkg7","Cxcr6" ,"Csf1","Rel"  , "Btg1","Ifngr1", "Il7r","Icos", "Ccr2","Ccl5","Fus","Gzmb","Ifit3","Slfn8","Ifit1",  "Ifi203","Isg15" ,"Gbp7" ,"Ly6a","Il1r2" ))+RotatedAxis()+theme(axis.title=element_blank())
```

```{r , fig.height =4, fig.width =12}
# --- STEP 20E: Refined Activated Treg Profiling ---
# A deep-dive DotPlot profiles the specific metabolic (Hilpda, Eno1), 
# translational (perc.ribo), and stress-related (Bmyc, Tff1) markers that 
# define the activation spectrum within the Treg compartment.
DotPlot(seur, features = c("Il1r2","Mrpl32","Pet100","Batf","Dgat1","Jarid2","Ptbp1","Il2rg","Bmp2k","perc.ribo","Rpl17","Fau",  "Hilpda" ,"Chchd10","Ppil3","Vim","Bmyc" , "Ddt","Dars","Slc39a7","Dnajb14","Tff1","Ripk3","Hmgn5","Scnm1", "Zfp593","Nhp2","Mif","Ran","Mki67","Stmn1","Ccr8","Eno1" ,"Bnip3" ,"doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```

```{r , fig.height =3.5, fig.width =5}
DotPlot(subset,features = c( "Il1r2","Ccr8"))+RotatedAxis()+theme(axis.title = element_blank())
```

## 21. Reciprocal Validation of the Bmyc-hi Gene Module

```{r}
# --- STEP 21A: Signature Extraction and Multi-Tiered Projection ---
# We extract the 'Bmyc-hi' signature from the MC38 reference (selected.genes) 
# and project it onto our current subset at three levels: 
# 1. Broadly by 'annot.general' (Resting vs. Activated).
# 2. Granularly by 'annot.detailed'.
# 3. Externally by the original MC38 Treg clusters.
# This confirms the signature's robustness across different clustering resolutions.
i="Treg.Il1r2.hi.Bmyc.hi"
selected.genes= DEgenes_MC38_Tregs [[i]][ DEgenes_MC38_Tregs[[i]]$p_val<0.01 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.35,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot.general", features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot.detailed", features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```


```{r , fig.height =3, fig.width =12}
DotPlot(MC38_Tregs, features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =10, fig.width =15}
# --- STEP 21B: Visualizing the Bmyc-hi Core Program ---
# We map the primary drivers of this subset—including NF-kB family members (Rel, Relb, Nfkb2), 
# activation markers (Cd83, Icam1), and stress/metabolic genes (Bmyc, Chchd10)—using 
# FeaturePlots. This spatial visualization in both the MC38 reference and the 
# experimental subset reveals a highly localized "hotspot" of these genes.
FeaturePlot(MC38_Tregs, features = c("Il1r2","Bmyc","Rel","Relb","Cd81","Bcl2a1b","Gpr183", "Gadd45b", "Chchd10","Penk" ,"Pcsk1" , "Ccdc184" , "Cd83" , "Bhlhe40" , "Icam1" ,"Nfkb2","Nrip1","Fas" ,"Marcksl1","Hspa9"  ),ncol=5, reduction ="umapHarmonyTheta1PC25")+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =4, fig.width = 4}
DimPlot(MC38_Tregs,cells.highlight = WhichCells(MC38_Tregs, idents="Treg Il1r2-hi Bmyc-hi"), sizes.highlight = 0.3,reduction ="umapHarmonyTheta1PC25")+NoLegend()
```

```{r , fig.height =4, fig.width = 6}
DimPlot(MC38_Tregs,label=T,repel=T, cols=cols.use,reduction ="umapHarmonyTheta1PC25")
```
```{r , fig.height =2.5, fig.width =10}
# --- STEP 21C: Comparative Signature "Fingerprinting" ---
# Using DotPlots to compare the expression intensity of the Bmyc module. 
# The patterns in the MC38 reference (Step 21B/C) nearly perfectly mirror 
# the experimental subset (Step 21C/D), proving that the 'Bmyc-hi' state is 
# a distinct and reproducible biological entity.
DotPlot(MC38_Tregs, features = c("Il1r2","Bmyc","Rel","Relb","Cd81","Bcl2a1b","Gpr183", "Gadd45b", "Chchd10","Penk" ,"Pcsk1" , "Ccdc184" , "Cd83" , "Bhlhe40" , "Icam1" ,"Nfkb2","Nrip1","Fas" ,"Marcksl1","Hspa9"    ))+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =10, fig.width =15}
FeaturePlot(subset, c("Il1r2","Bmyc","Rel","Relb","Cd81","Bcl2a1b","Gpr183", "Gadd45b", "Chchd10","Penk" ,"Pcsk1" , "Ccdc184" , "Cd83" , "Bhlhe40" , "Icam1" ,"Nfkb2","Nrip1","Fas" ,"Marcksl1","Hspa9"   ),ncol=5,reduction = paste0("umapPC",dims.use.subset))+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =2.5, fig.width =10}
DotPlot(subset, group.by = "annot.general", features = c("Il1r2","Bmyc","Rel","Relb","Cd81","Bcl2a1b","Gpr183", "Gadd45b", "Chchd10","Penk" ,"Pcsk1" , "Ccdc184" , "Cd83" , "Bhlhe40" , "Icam1" ,"Nfkb2","Nrip1","Fas" ,"Marcksl1","Hspa9"    ))+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =4, fig.width = 4}
# --- STEP 21D: Final Cluster Localization and Color Mapping ---
# We use 'cells.highlight' to confirm the spatial position of the Bmyc-hi 
# population. Finally, a customized DimPlot is generated with specific colors 
# to highlight the "Activated" Treg flavors (IFN-hi in orchid, Bmyc-hi in magenta, 
# and Proliferative in goldenrod) against the grey resting Treg background.
DimPlot(subset,cells.highlight = WhichCells(subset, idents="Treg Il1r2-hi Bmyc-hi"), sizes.highlight = 0.8,reduction = paste0("umapPC",dims.use.subset))+NoLegend()
```


```{r , fig.height =4, fig.width = 6}
DimPlot(subset, group.by = "annot.general",cols=c("Treg"="grey"   , "Treg IFN"="grey" ,   "Treg Il1r2-hi IFN" ="darkorchid",  "Treg Il1r2-hi" ="darkolivegreen1" ,  "Treg Il1r2-hi Bmyc-hi"="magenta" ,"Treg Il1r2-hi prolif"="goldenrod1" ), sizes.highlight = 0.8,reduction = paste0("umapPC",dims.use.subset))
```

## 22. Ccr8-hi Signature Projection and Final Data Serialization

```{r}
# --- STEP 22A: Ccr8-hi Effector Program Projection ---
# We extract the 'Ccr8-hi' signature from the MC38 reference and map it onto 
# our experimental subset. This identifies the most potent effector pole of 
# the Treg population, characterized by high expression of suppressive 
# molecules (Il10) and chemotactic factors (Ccl4).
i= "Treg.Il1r2.hi.Ccr8.hi"
selected.genes= DEgenes_MC38_Tregs[[i]][ DEgenes_MC38_Tregs[[i]]$p_val<0.01 &  DEgenes_MC38_Tregs[[i]]$avg_logFC>0.3,"gene"]
```


```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot.general",features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =3, fig.width =12}
DotPlot(subset, group.by = "annot.detailed", features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =3, fig.width =12}
DotPlot(MC38_Tregs, features = unique(c("Il1r2",selected.genes)))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =8, fig.width =12}
# --- STEP 22B: Spatial Mapping of Effector Markers ---
# FeaturePlots visualize the expression of Ccr8, Il10, and Vsir (VISTA) in 
# both datasets. The co-localization of these genes confirms that the 
# Il1r2-hi Bmyc-hi/prolif clusters overlap with the most immunosuppressive 
# "terminal" effector state identified in the MC38 tumor model.
FeaturePlot(MC38_Tregs, features = c("Il1r2","Ccr8",  "Ccl4" , "Rilpl2" , "Mt1",  "Sv2c"  ,"Il10",  "Bcl2a1d", "Vsir","Plk3","Eea1","Irf8"   ), reduction ="umapHarmonyTheta1PC25")+RotatedAxis()+theme(axis.title = element_blank())
```
```{r , fig.height =4, fig.width = 6}
DimPlot(MC38_Tregs,label=T,repel=T, cols=cols.use,reduction ="umapHarmonyTheta1PC25")
```

```{r , fig.height =8, fig.width =12}
FeaturePlot(subset, features = c("Il1r2","Ccr8",  "Ccl4" , "Rilpl2" , "Mt1",  "Sv2c"  ,"Il10",  "Bcl2a1d", "Vsir","Plk3","Eea1","Irf8"   ) ,reduction = paste0("umapPC",dims.use.subset))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height = 3.8, fig.width =14}
DotPlot(MC38_Tregs,features=c("Cd4","Foxp3","Ikzf2",  "Ctla4","Saraf","Klf2","Klf3","Itgb1","Btg2","Ier2","Fos","Lef1","Cd7"   ,"Tnfrsf9","Il2ra", "Arl5a","Ctla2a","Socs2","Eno1" ,"Cxcr3","Bnip3" ,"Ifit3","Ifit2", "Isg15","Sell"   ,"Lag3","Pdcd1","Tigit","Ccl5","Nkg7","Il1r2","Ccr8" ,"Cst7" ,"Bcl2a1b",  "Ccl4"  , "Rilpl2" ,"Bcl2a1d","Il10","Vsir","Irf8"    ,"Bmyc","Nrip1", "Chchd10","Bhlhe40" ,"Icam1","Nfkbia","Ran","Mki67","Stmn1"))+RotatedAxis()
```



```{r , fig.height =4, fig.width =15}
DotPlot(subset, features = c( "Foxp3","Il2ra","Ctla4","Cst7" ,"Bcl2a1b","Nkg7","Cxcr6" ,"Csf1","Rel"  , "Btg1","Ifngr1", "Il7r","Icos", "Ccr2","Ccl5","Fus","Gzmb","Ifit3","Slfn8","Ifit1",  "Ifi203","Isg15" ,"Gbp7" ,"Ly6a","Il1r2","Mrpl32","Pet100","Batf","Dgat1","Jarid2","Ptbp1","Il2rg","Bmp2k","perc.ribo","Rpl17","Fau",  "Hilpda" ,"Chchd10","Ppil3","Vim","Bmyc" , "Ddt","Dars","Slc39a7","Dnajb14","Tff1","Ripk3","Hmgn5","Scnm1", "Zfp593","Nhp2","Mif","Ran","Mki67","Stmn1","Ccr8","Eno1" ,"Bnip3" ,"doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```

```{r , fig.height =3.5, fig.width =5}
# --- STEP 22C: Broad vs. Detailed Marker Comparison ---
# A comprehensive DotPlot analysis confirms the expression of classic Treg 
# markers (Foxp3, Ctla4) alongside activation-induced genes (Eno1, Bnip3). 
# This confirms that 'Il1r2' is a reliable proxy for the transition from 
# resting (Sell-hi/Klf2-hi) to activated (Ccr8-hi/Il1r2-hi) states.
DotPlot(subset,  group.by = "annot.general",features = c( "Il1r2","Ccr8"))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =4, fig.width =6,warning=FALSE}
DimPlot(subset, group.by = "annot.general",label=T,repel=T,cols= cols.use,reduction = paste0("umapPC",dims.use.subset)) 
```

```{r , fig.height =4, fig.width =12}
DotPlot(subset, group.by = "annot.general", features = c( "Foxp3","Il2ra","Ctla4","Ccl5","Ly6e","Ifit3","Slfn8","Ifit1",  "Ifi203","Isg15" ,"Gbp7" ,"Ly6a","Il1r2","Pet100","Batf","Dgat1","perc.ribo","Rpl17","Fau", "Ddit4", "Hilpda" ,"Chchd10","Ppil3","Vim","Bmyc" , "Ddt","Polr3c","Dars","Slc39a7","Dnajb14","Tff1","Ripk3","Hmgn5","Scnm1", "Zfp593","Nhp2","Mif","Ran","Ccr8","Eno1"  ,"doublet.score" ,"perc.mito","nCount_RNA_log","nFeature_RNA_log"))+RotatedAxis()
```

```{r , fig.height =3.7, fig.width =12}
DotPlot(subset, group.by = "annot.general", features = c( "Foxp3","Il2ra","Ctla4","Ccl5","Ly6e","Ifit3","Slfn8","Ifit1",  "Ifi203","Isg15" ,"Gbp7" ,"Ly6a","Il1r2","Pet100","Batf","Dgat1","perc.ribo","Rpl17","Fau",  "Hilpda" ,"Chchd10","Ppil3","Vim","Bmyc" , "Ddt","Dars","Slc39a7","Dnajb14","Tff1","Ripk3","Hmgn5","Scnm1","Polr3c","Nhp2","Mif","Ran","Mki67","Stmn1"))+RotatedAxis()+theme(axis.title = element_blank())
```

```{r , fig.height =4, fig.width =4.3,warning=FALSE, echo=FALSE}
FeaturePlot(object = subset,reduction = paste0("umapPC",dims.use.subset), c("Il1r2"))
```

```{r , fig.height =3, fig.width =5}
DotPlot(subset, features = c( "Il1r2"))+RotatedAxis()
```


```{r , fig.height =5, fig.width =5}
# --- STEP 22D: Quantifying Activation via Il1r2 ---
# Violin plots and FeaturePlots of 'Il1r2' provide a final quantitative 
# look at the activation gradient. This clearly separates the 'General' 
# Treg categories, proving that Il1r2 expression strictly follows the 
# transition to the terminal effector state.
VlnPlot(subset, features = c( "Il1r2"), group.by = "annot.general", cols=cols.use)+RotatedAxis()+NoLegend()+theme(axis.title.x=element_blank())
```


```{r}
# --- STEP 22E: Final Data Serialization and Archiving ---
# The finalized, annotated Seurat objects—including the full dataset, 
# the T/NK subset, and the high-resolution Treg subset—are saved as .rds files. 
# This ensures that all metadata, cluster labels, and dim-reductions are 
# preserved for downstream differential abundance or pseudotime analysis.
saveRDS(subset, paste0(path.output,subset_name,"_",dataset.name,".seurat.rds"))
```

```{r}
subset=readRDS( paste0(path.output,subset_name,"_",dataset.name,".seurat.rds"))
dims.use.subset=25
```
```{r}
subset
```



```{r}
saveRDS(seur, paste0(path.output,"T_NK_subset_",dataset.name,".seurat.rds"))
```

```{r}
seur=readRDS( paste0(path.output,"T_NK_subset_",dataset.name,".seurat.rds"))
```
```{r}
seur
```

```{r}
seur_full
```
```{r,warning=FALSE}
sessionInfo()
```




